### 2020-09-06 HW1/Regression Model - Deokinanan.rmd
Added regression model #1

---
title: "DATA 621 - HW1 - Regression Model"
author: "Samantha Deokinanan"
date: "September 3, 2020"
output: html_document
---

***
#### Task
***

*Using the training data set, build at least three different multiple linear regression models, using different variables(or the same variables with different transformations). Since we have not yet covered automated variable selection methods, you should select the variables manually (unless you previously learned Forward or Stepwise selection, etc.). Since you manually selected a variable for inclusion into the model or exclusion into the model, indicate why this was done.*

*Discuss the coefficients in the models, do they make sense? For example, if a team hits a lot of Home Runs, it would be reasonably expected that such a team would win more games. However, if the coefficient is negative (suggesting that the team would lose more games), then that needs to be discussed. Are you keeping the model even though it is counter intuitive? Why? The boss needs to know.*

***
#### The dataset
***

The data set contains approximately 2200 records. Each record represents a professional baseball team from the years 1871 to 2006 inclusive. Each record has the performance of the team for the given year, with all of the statistics adjusted to match the performance of a 162 game season. Aim is to predict the number of wins for the team.

```{r message=FALSE, warning=FALSE}
# The required R packages
library(tidyverse)
library(psych)
library(olsrr)
library(rcompanion)
library(mice)
library(VIM)
library(mctest)
library(caret)

# Load training data set
train_df = read.csv("C:/Users/Deokinanan/Desktop/2019 -2021 CUNY SPS/FALL 2020/DATA 621/WEEK 3/moneyball-training-data.csv")
train_df = subset(train_df, select = -c(INDEX))
```

***
#### Data Transformation, Outliers and Missing Data
***

A quick exploration of the data, and it is already apparent that the `TEAM_BATTING_HBP` variable has nearly 92% of the its data missing. Since it would be very difficult to accurately impute such a large proportion, this variable will be excluded from further analysis.

```{r}
# Quick Exploration
sapply(train_df, describe)

# Data Transformation
train_df = subset(train_df, select = -c(TEAM_BATTING_HBP))
```

Next, the outlier plots revealed that there are a few extreme values that can influence the analysis. Because the objective is so create a multivariate regression model, declaring an observation as an outlier based on a just one feature could lead to unrealistic inferences. There Cook's distance is use to decide if an individual entity is an extreme value or not.

Cook’s distance is a measure computed with respect to a given regression model and therefore is impacted only by the predictors included in the model. It computes the influence exerted by each data point on the response variable. Fitting the full model during this exploratory stage, the process remove the most influential points. Note from the plots that all influential observations are not necessarily outliers. Least than 3% of the data was removed.

```{r}
# Outlier Plots
model = lm(TARGET_WINS ~ ., data = train_df)
ols_plot_diagnostics(model)

# Remove Outlier
cooksd = cooks.distance(model)
influential = as.numeric(names(cooksd)[(cooksd > 4*mean(cooksd, na.rm = TRUE))])
train_df = train_df[-c(influential), ]

# New diagnostic plots
model_new = lm(TARGET_WINS ~ ., data = train_df)
ols_plot_diagnostics(model_new)
```

Now, an investigation is perform to calculate the amount of missing/imputed values in each variable. The plot helps to understand that almost 85% of the samples are not missing any information, whereas below states the percentage of missing values per variables. The method of choice to handle these missing values is imputation using predictive mean matching, to replace missing data with a randomly chosen value from several similar cases.

```{r}
aggr(train_df, col = c('darkseagreen2','darkslateblue'), numbers = TRUE, sortVars = TRUE, oma = c(10,5,5,3), labels = names(train_df), cex.axis = 0.8, gap = 3, axes = TRUE, Prop = TRUE, ylab = c("Proportion of Missing Data", "Combination"))

train_df.clean = complete(mice(data = train_df, method = "pmm", maxit = 5, seed = 525, print = FALSE), 3)
sapply(train_df.clean, describe)
```

***
#### Initial Tests 
***

Because the data consist of variable called `TEAM_BATTING_H`, base hits by batters, i.e. it is linear combination of H = 1B + 2B + 3B + HR, there are concerns of possible multicollinearity. Therefore, the single hits by batter found and base hits were removed. Next, a collinearity diagnostic test is done to examining the diagnostic output for variance inflation factor, tolerance, and Farrar-Glauber F-test. The F-statistic for the variable `TEAM_BATTING_HR` is quite high (42.1158) followed by the variable `TEAM_PITCHING_HR` (F-value of 33.8885). So, the test shows that there are multiple variables that will be the root cause of multicollinearity. Moreover, as expected, there are high partial correlations found to be statistically significant. As a solution to deal with multicollinearity, there are several remedial measures will be used as a result of this diagnostic test. Some included removal of highly correlated variables and stepwise regression analysis were done.

```{r}
# Data Transformation
train_df.clean$TEAM_BATTING_1B = train_df.clean$TEAM_BATTING_H - (train_df.clean$TEAM_BATTING_2B + train_df.clean$TEAM_BATTING_3B + train_df.clean$TEAM_BATTING_HR)
train_df.clean = subset(train_df.clean, select = -c(TEAM_BATTING_H))

# Multicollinearity Diagnostic Measures
model_new = lm(TARGET_WINS ~ ., data = train_df.clean)
imcdiag(model_new, method = "VIF", all = TRUE)

# Data Transformation & Multicollinearity Diagnostic Measures
train_df.clean = subset(train_df.clean, select = -c(TEAM_PITCHING_HR))
model_new = lm(TARGET_WINS ~ ., data = train_df.clean)
imcdiag(model_new, method = "VIF", all = TRUE)
```

```{r, fig.width=7, fig.height=9}
# Correlation
corr = round(cor(train_df.clean), 2)
corr = cbind(names = rownames(corr), data.frame(corr, row.names=NULL))
g.corr = gather(as.data.frame(corr), key, value, 2:length(corr), na.rm = TRUE)

ggplot(data = g.corr, aes(key, names, fill = value)) + geom_tile(color = "white") +
 scale_fill_gradient2(low = "darkseagreen2", high = "darkslateblue", mid = "white", 
   midpoint = 0, limit = c(-1,1), space = "Lab", name="Pearson\nCorrelation") +
  theme_minimal() + theme(axis.text.x = element_text(angle = 45, vjust = 1, size = 12, hjust = 1)) +
 coord_fixed() + geom_text(aes(key, names, label = value), color = "black", size = 3) 
```

Lastly, because a robust regression process will be performed, higher order polynomials variables were introduced into the full model.

```{r}
stepdata = cbind(train_df.clean,sapply(train_df.clean[2:length(train_df.clean)], function(x) x^2))
names(stepdata) = make.unique(names(stepdata), sep = "_")
```

***
#### Regression Model #1
***

A stepwise variable selection model is conducted to determine what are the variables that can help predict the number of wins for the team. The stepwise variable selection allows variables to be added one at a time to the model, as long as the F-statistic is below the specified $\alpha$, in this case $\alpha = 0.05$. However, variables already in the model do not necessarily stay in. The steps evaluate all of the variables already included in the model and remove any variable that has an insignificant F-statistic. Only after this test ends, is the best model found, that is when none of the variables cstepan be excluded and every variable included in the model is significant.

Here the dependent variable is the continuous variable, `TARGET_WINS`, and the independent variables are the full model to identify the most contributing predictors. In addition, a robust method for estimating the accuracy of a model, the k-fold cross-validation method, was performed evaluate the model performance on different subset of the training data and then calculate the average prediction error rate.

After the steps, the final model resulted below, with $adj. R^2 = 0.42$, suggesting that this model accounts for nearly 42% of the variation in the dependent variable with the independent variables, which is acceptable as a good model. Within the method of `lmStepAIC`, AIC (Akaike Information Criteria) quantifies the amount of information loss due to simplification. That is, based on the AIC, the final model outputted is the simplest model without impacting much on the performance.

```{r  message=FALSE, warning=FALSE, paged.print=FALSE}
# Set up repeated k-fold cross-validation
set.seed(525)
train.control = trainControl(method = "cv", number = 10)

# Train the model
step.model = train(TARGET_WINS ~ ., data = stepdata, method = "lmStepAIC", trControl = train.control, trace = FALSE)

# Model accuracy
step.model$results

# Final model coefficients
step.model$finalModel

# Summary of the model
summary(step.model$finalModel)
```

***
#### Prediction
***

Studying the coefficients of the model suggest that winning is in favor if the team batting hits more doubles, triples and home runs. Moreover, increase in the number of stolen bases, and a decrease in caught steals, double plays, error, and walks allowed would all lead to a win for the batting team. It is noteworthy that the model suggests that a decrease in single hits by batter and an increase in strikeouts by batters which seems counter intuitive. But these variables were kept because when a batter steps to the plate, the player is more likely to strike out than to get a hit. Trying to hit the ball out of the park will come with strikeouts but it will also increase the chances of hitting home runs (even 1B, 2B, 3B), and that is pretty good exchange that most teams are willing carry out.

```{r message=FALSE, warning=FALSE}
# load the test data
test_df = read.csv("C:/Users/Deokinanan/Desktop/2019 -2021 CUNY SPS/FALL 2020/DATA 621/WEEK 3/moneyball-evaluation-data.csv")

# Data Transformation to match train data
test_df = complete(mice(data = test_df, method = "pmm", maxit = 5, seed = 525, print = FALSE), 3)
test_df$TEAM_BATTING_1B = test_df$TEAM_BATTING_H - (test_df$TEAM_BATTING_2B + test_df$TEAM_BATTING_3B + test_df$TEAM_BATTING_HR)
test_df = cbind(test_df,sapply(test_df[1:length(test_df)], function(x) x^2))
names(test_df) = make.unique(names(test_df), sep = "_")

predictions = predict(step.model, newdata = test_df)

results = cbind(dataset = c('Training Data', 'Test Prediction'), bind_rows(describe(train_df$TARGET_WINS), describe(predictions))) 
rownames(results) = NULL

results
```