### 2020-09-08 HW1/RegModel_Adler.Rmd
Imputation, center/scaling, start commentary on results of c/s

---
title: "Data 621---HW1---Regression Model---Version 2"
author: "Group 2"
date: "9/8/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```
```{r loadPackages}
library(caret)
library(corrplot)
library(RColorBrewer)
library(ggplot2)
library(scales)
library(data.table)
```

# Data Exploration
## Task
Describe the size and the variables in the moneyballtraining data set. Consider
that too much detail will cause a manager to lose interest while too little
detail will make the manager consider that you aren’t doing your job. Some
suggestions are given below. Please do NOT treat this as a check list of things
to do to complete the assignment. You should have your own thoughts on what to
tell the boss. These are just ideas.

  a. Mean / Standard Deviation / Median
  b. Bar Chart or Box Plot of the data
  c. Is the data correlated to the target variable (or to other variables?)
  d. Are any of the variables missing and need to be imputed “fixed”?

## Analysis
```{r readData}
MB <- fread('./data/moneyball-training-data.csv')
n <- dim(MB)[[1]]
k <- dim(MB)[[2]]

# Meltng data from wide to narrow for munging & ggplot; simultaneously removing
# INDEX.
MBmelt <- melt(MB[, -1L], measure.vars = names(MB[, -1L]),
               variable.name = 'Metric', value.name = 'Value',
               variable.factor = FALSE)
```

### Description
The data is composed of `r n` observations of `r k` fields, all of which are
discrete numerical variables.

### Missing Observations
The table below shows all variables with missing values:

```{r missingTable}
missingObs <- MBmelt[, .(Missing = sum(is.na(Value)),
                         Percentage = sum(is.na(Value)) / length(Value)),
                     keyby = Metric][Missing > 0][order(-Missing)]
knitr::kable(missingObs, digits = 3L,
caption = 'Training Data with Missing Observations')
```

Most variables are present in their entirety. Some have a few observations
missing that may be addressed through imputation. Unfortunately,
`TEAM_BATTING_HBP`---the metric representing batters getting first base awarded
due to being hit by a pitch---is missing almost all observations. It is prudent
to remove that variable from the analysis. It is also probably prudent to
remove `TEAM_BASERUN_CS` as it is missing more than one-third of the time. The
variables missing less than 10% of the time may be imputed using median or
k-means clustering. Whether `TEAM_FIELDING_DP` will be removed or imputed will
be a decision made later on in the analysis.

### Data Distribution
The table and plots below give an overview as to the distribution of the
variables.

```{r summaryTable}
# Removing 'TEAM_BATTING_HBP' and index from wide data. Easier to calculate
# correlations later on using wide data.
MB[, `:=`(INDEX = NULL, TEAM_BATTING_HBP = NULL)]

# Removing 'TEAM_BATTING_HBP' from tall data
MBmelt <- MBmelt[Metric != 'TEAM_BATTING_HBP']

# Casting to double since medians and quartiles may return non-integer values.
MBmelt[, Value := as.double(Value)]

statTable <- MBmelt[, .(Min = min(Value, na.rm = TRUE),
                        Q1 = quantile(Value, 0.25, na.rm = TRUE),
                        Median = median(Value, na.rm = TRUE),
                        Q3 = round(quantile(Value, 0.75, na.rm = TRUE)),
                        Max = max(Value, na.rm = TRUE),
                        Mean = mean(Value, na.rm = TRUE),
                        StDev = sd(Value, na.rm = TRUE),
                        IQR = IQR(Value, na.rm = TRUE)), keyby = Metric]
knitr::kable(statTable, digits = 2L,
             caption = 'Training Data Summary Statistics')
```

```{r boxPlot1, fig.width = 7}
ggplot(MBmelt, aes(x = Value, y = Metric)) + geom_boxplot() +
  scale_x_continuous(labels = comma)
```

One immediate observation from both the table and the boxplots, is that the
range for `TEAM_PITCHING_H` is one to two orders of magnitude greater than that
of the other variables. Below are boxplots of the remaining variables for a
better view, followed by a histogram of `TEAM_PITCHING_H`.

```{r boxPlot2, fig.width = 7}
ggplot(MBmelt[Metric != 'TEAM_PITCHING_H'], aes(x = Value, y = Metric)) +
  geom_boxplot() + scale_x_continuous(labels = comma)

# Use the Freedman-Diaconis rule for historgram bin widths instead of ggplot
# defaults

FreedmanDiaconis <- function(x) {
  2 * IQR(x, na.rm = TRUE) / (length(x) ^ (1 / 3))
}

ggplot(MBmelt[Metric == 'TEAM_PITCHING_H'], aes(x = Value)) +
  geom_histogram(binwidth = FreedmanDiaconis) +
  scale_x_continuous(labels = comma)
```

It is very clear that `TEAM_PITCHING_H` is extremely right tailed and may be a
prime subject for a transformation.

### Correlations
A quick way to identify correlations between variables is through a
corrgram.[^1] In the corrgram below, blue represents positive correlation and
red represents negative correlation, as shown in the color bar. The shape of the
images in each cell reflect the scatterplots of the features with each other.
The main diagonal, therefore, are all upward sloping blue lines, as each feature
is perfectly correlated with itself.

[^1]:This is the proper name for a multicolored matrix representation of a
correlation matrix (Friendly 2002). Correlelograms refer to the time-series
plots used to check for auto-correlation. The terms have been conflated, though.

```{r corrgram, fig.height=6}
MBcor <- cor(MB, method = 'pearson', use = 'pairwise.complete.obs')
corrplot(MBcor, 'ellipse', type = 'lower', order = 'hclust',
         col=brewer.pal(n=8, name="RdYlBu"))
```

Most of the ellipses are pretty wide, so there is not *much* correlation, but
there are some clear clusters. For example, `TEAM_FIELDING_E` and 
`TEAM_BATTING_3B` are both negatively correlated with `TEAM_BATTING_SO` and
`TEAM_BATTING_HR`. As regards `TEAM_FIELDING_E`, a hypothesis is that if batters
are either hitting home runs or striking out, *they are***not***getting on base*.
Therefore there are fewer opportunity for fielding errors. Why triples are 
negatively correlated with HRs or SOs is not as logical. There is a positive
correlation between triples and fielding errors of around 51%. Perhaps triples
are more a function of fielding than batting?

Regardless, there are some correlations in the data, but nothing that screams
for immediate action.

# Data Preperation
## Task
Describe how you have transformed the data by changing the original variables or
creating new variables. If you did transform the data or create new variables,
discuss why you did this. Here are some possible transformations:

  a. Fix missing values (maybe with a Mean or Median value)
  b. Create flags to suggest if a variable was missing
  c. Transform data by putting it into buckets
  d. Mathematical transforms such as log or square root(or use Box-Cox)
  e. Combine variables (such as ratios or adding or multiplying) to create new
  variables

## Analysis
The `caret` package for R is the gold-standard catch-all interface for machine
learning techniques for those not using the tidyverse. Otherwise, the same
author's `tidymodels` package may be used as well. When using `caret`, almost
all of the following imputation and transformation is done as pre-processing
to the model. For the purposes of this homework, they are being performed
manually to demonstrate understanding.

### Imputation
Since we are restricting ourselves to the world of the simple linear regression
model, it makes sense to use a Z-transformation to standardize the models.
Prior to this, the missing variables should be imputed. Looking at the summary
statistics table, the median is not that far off of the mean in most cases with
the exceptions of `TEAM_BASERUN_SB` and `TEAM_BATTING_SO`. Therefore, for an
initial imputation, we will use the mean so that the overall mean is unaffected
for the Z-transform. This **will** artificially reduce the SD, however. We could
address that by calculating the mean and SD, *prior* to imputation. We will not
do so now for simplicity.

```{r imputeMean}
MBmelt[, Value := replace(Value, is.na(Value), mean(Value, na.rm = TRUE)),
       keyby = Metric]
```

### Transformations
The first transform applied will be centering and scaling, as discussed above.

```{r centerScale}
MBmelt[, Value := scale(Value, center = TRUE, scale = TRUE), keyby = Metric]
```

At this point we would hope to see Gaussian-looking densities.
```{r plotScaled}
ggplot(MBmelt, aes(x = Value)) + geom_density() +
  facet_wrap(~ Metric, scales = 'free')
```

Some of the variables, like `TEAM_BATTING_2B` or `TEAM_BATTING_SO` show
Gaussian-like behavior, with some allowances for the discrete nature of the
input data (those wiggles at the crest of `TEAM_BATTING_SO` for example). Others
are classic right tailed (exponential/gamma-esque) such as `TEAM_FIELDING_E` or
`TEAM_BATTING_3B`. Still others are clearly multimodal. Then there are the
weird spike-like ones as well.

# References

 * Friendly, Michael (2002) Corrgrams: Exploratory displays for correlation
 matrices, *The American Statistician*, 56:4, 316-324, DOI: 10.1198/000313002533,
 http://euclid.psych.yorku.ca/datavis/papers/corrgram.pdf

# Appendix
The code chunks below represent the code called in order during the analysis.
They are reproduced in the appendix for review and comment.

```{r loadPackages, echo=TRUE, eval=FALSE}
```
```{r readData, echo=TRUE, eval=FALSE}
```
```{r missingTable, echo=TRUE, eval=FALSE}
```
```{r summaryTable, echo=TRUE, eval=FALSE}
```
```{r boxPlot1, echo=TRUE, eval=FALSE}
```
```{r boxPlot2, echo=TRUE, eval=FALSE}
```
```{r corrgram, echo=TRUE, eval=FALSE}
```
```{r imputeMean, echo=TRUE, eval=FALSE}
```
```{r centerScale, echo=TRUE, eval=FALSE}
```