### 2020-09-09 HW1/HW1_AF.Rmd
data exploration - initial

---
title: 'DATA 621---Business Analytics and Data Mining'
subtitle: 'Fall 2020---Group 2---Homework #1'
author: Avraham Adler, Samantha Deokinanan, Amber Ferger, John Kellogg,
    Bryan Persaud, Jeff Shamp
date: "9/27/2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r loadData, include=FALSE}
library(tidyverse)
library(ggplot2)
library(corrplot)
train_set <- read.csv('data/moneyball-training-data.csv')
eval_set <- read.csv('data/moneyball-evaluation-data.csv')
```

## 1. DATA EXPLORATION (Amber)
Describe the size and the variables in the moneyball training data set. Consider
that too much detail will cause a manager to lose interest while too little
detail will make the manager consider that you aren't doing your job. Some
suggestions are given below. Please do NOT treat this as a check list of things
to do to complete the assignment. You should have your own thoughts on what to
tell the boss. These are just ideas.

  a. Mean / Standard Deviation / Median
  b. Bar Chart or Box Plot of the data
  c. Is the data correlated to the target variable (or to other variables?)
  d. Are any of the variables missing and need to be imputed “fixed”?

We are given a data set containing 2,276 records, each of which includes the 
number of wins the team had that year and 15 additional statistics about the 
season. A first look at the data shows that only about 8% of the records have 
a full set of data. The good news is that most of the missing values come from 
what we would expect; hit-by-pitch, caught stealing, double plays.

```{r, echo = FALSE, message = FALSE, warning = FALSE}

pct_null <- data.frame(do.call("rbind", map(train_set %>% select(-INDEX), ~mean(is.na(.)))))
colnames(pct_null) <- c('PCT_NULL')

pct_null %>%
  arrange(desc(PCT_NULL)) %>%
  filter(PCT_NULL > 0)

```

  * Total cases: 2276, Complete cases: 191 (only ~8% of the data has fully filled in values)
  * 15 stats (variables), 1 target (wins)
  * Best team won 146 out of 162 games, worst team won none, Average number of wins is 80 (show win dist)
  * Discuss missing values

```{r, echo = FALSE, message = FALSE, warning = FALSE}

# histogram of wins
p <- ggplot(train_set, aes(x=TARGET_WINS)) + 
  geom_histogram() + 
  geom_vline(aes(xintercept=mean(TARGET_WINS)),
            color="blue", linetype="dashed", size=1)+
  labs(title="Distribution of Wins",x="Number of wins", y = "Number of teams")

p

# Correlation matrix with significance levels (p-value)
library("Hmisc")
res2 <- rcorr(as.matrix(train_set %>% select(-INDEX)))

# Insignificant correlation are crossed
corrplot(res2$r, type="upper", order="hclust", 
          tl.col = "black", p.mat = res2$P, sig.level = 0.01, insig = "blank")

train_set %>%
  select(-INDEX) %>%
  gather(-TARGET_WINS, key = "STAT", value = "VALUE") %>%
  ggplot(aes(x = VALUE, y = TARGET_WINS)) +
    geom_point() +
  labs(title="Number of wins with respect to recorded stats",x="", y = "Number of Wins") +
    facet_wrap(~ STAT, scales = "free")

```

## 2. DATA PREPARATION (25 Points)
Describe how you have transformed the data by changing the original variables or
creating new variables. If you did transform the data or create new variables,
discuss why you did this. Here are some possible transformations.

  a. Fix missing values (maybe with a Mean or Median value)
  b. Create flags to suggest if a variable was missing
  c. Transform data by putting it into buckets
  d. Mathematical transforms such as log or square root (or use Box-Cox)
  e. Combine variables (such as ratios or adding or multiplying) to create new
  variables

## 3. BUILD MODELS (25 Points)
Using the training data set, build at least three different multiple linear
regression models, using different variables (or the same variables with
different transformations). Since we have not yet covered automated variable
selection methods, you should select the variables manually (unless you
previously learned Forward or Stepwise selection, etc.). Since you manually
selected a variable for inclusion into the model or exclusion into the model,
indicate why this was done.

Discuss the coefficients in the models, do they make sense? For example, if a
team hits a lot of Home Runs, it would be reasonably expected that such a team
would win more games. However, if the coefficient is negative (suggesting that
the team would lose more games), then that needs to be discussed. Are you
keeping the model even though it is counter intuitive? Why? The boss needs to
know.

## 4. SELECT MODELS (25 Points)
Decide on the criteria for selecting the best multiple linear regression model.
Will you select a model with slightly worse performance if it makes more sense
or is more parsimonious? Discuss why you selected your model.

For the multiple linear regression model, will you use a metric such as Adjusted
\(R^2\), RMSE, etc.? Be sure to explain how you can make inferences from the
model, discuss multi-collinearity issues (if any), and discuss other relevant
model output. Using the training data set, evaluate the multiple linear
regression model based on (a) mean squared error, (b) \(R^2\), (c) F-statistic,
and (d) residual plots. Make predictions using the evaluation data set.

## APPENDIX
### Assignment Overview
In this homework assignment, you will explore, analyze and model a data set
containing approximately 2200 records. Each record represents a professional
baseball team from the years 1871 to 2006 inclusive. Each record has the
performance of the team for the given year, with all of the statistics adjusted
to match the performance of a 162 game season.

Your objective is to build a multiple linear regression model on the training
data to predict the number of wins for the team. You can only use the variables
given to you (or variables that you derive from the variables provided). Below
is a short description of the variables of interest in the data set:

\begin{center}
\begin{tabular}{|l|l|l|}
\hline
\textbf{VARIABLE NAME} & \textbf{DEFINITION} & \textbf{THEORETICAL EFFECT} \\ \hline
INDEX & Identification Variable (do not use) & None \\ \hline
TARGET\_WINS & Number of wins & \\ \hline
TEAM\_BATTING\_H & Base Hits by batters (1B,2B,3B,HR) & Positive Impact on Wins  \\ \hline
TEAM\_BATTING\_2B & Doubles by batters (2B) & Positive Impact on Wins \\ \hline
TEAM\_BATTING\_3B & Triples by batters (3B) & Positive Impact on Wins \\ \hline
TEAM\_BATTING\_HR & Homeruns by batters (4B) & Positive Impact on Wins \\ \hline
TEAM\_BATTING\_BB & Walks by batters & Positive Impact on Wins \\ \hline
TEAM\_BATTING\_HBP & Batters hit by pitch (get a free base) & Positive Impact on Wins \\ \hline
TEAM\_BATTING\_SO & Strikeouts by batters & Negative Impact on Wins \\ \hline
TEAM\_BASERUN\_SB & Stolen bases & Positive Impact on Wins \\ \hline
TEAM\_BASERUN\_CS & Caught stealing & Negative Impact on Wins \\ \hline
TEAM\_FIELDING\_E & Errors & Negative Impact on Wins \\ \hline
TEAM\_FIELDING\_DP & Double Plays & Positive Impact on Wins \\ \hline
TEAM\_PITCHING\_BB & Walks allowed & Negative Impact on Wins \\ \hline
TEAM\_PITCHING\_H & Hits allowed & Negative Impact on Wins \\ \hline
TEAM\_PITCHING\_HR & Homeruns allowed & Negative Impact on Wins \\ \hline
TEAM\_PITCHING\_SO & Strikeouts by pitchers & Positive Impact on Wins \\ \hline

\end{tabular}
\end{center}

### R Programming code

```{r initial, warning = FALSE}
########### libraries 
library(tidyverse)

########### data
train_set <- read.csv('data/moneyball-training-data.csv')
eval_set <- read.csv('data/moneyball-evaluation-data.csv')

```

```{r dataExploration, results = 'hide'}
# initial summary of all values 
summary(train_set)
p <- ggplot(train_set, aes(x=TARGET_WINS)) + 
  geom_histogram() + 
  geom_vline(aes(xintercept=mean(TARGET_WINS)),
            color="blue", linetype="dashed", size=1)+
  labs(title="Distribution of Wins",x="Number of wins", y = "Number of teams")

# complete vs total cases
train_set %>%
  filter(complete.cases(.) == TRUE) %>%
  count() %>%
  rename("complete cases"= n) %>%
  bind_cols(
    train_set %>%
    select(TARGET_WINS) %>%
    count() %>%
    rename("target count" = n)
  )
```