### 2020-09-09 HW1/HW1_AF.Rmd
data exploration updates

---
title: 'DATA 621---Business Analytics and Data Mining'
subtitle: 'Fall 2020---Group 2---Homework #1'
author: Avraham Adler, Samantha Deokinanan, Amber Ferger, John Kellogg,
    Bryan Persaud, Jeff Shamp
date: "9/27/2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r loadData, include=FALSE}
library(tidyverse)
library(ggplot2)
library(corrplot)
train_set <- read.csv('data/moneyball-training-data.csv')
eval_set <- read.csv('data/moneyball-evaluation-data.csv')
```

## 1. DATA EXPLORATION

### How often does the team win?

We are given a data set containing 2,276 records, each of which includes the 
number of wins the team had that year and 15 additional statistics about the 
season. The team wins on average about 50% of the games they play (81 games 
out of 162), with the best season having 146 wins and the worst season having 
0 wins. The data is pretty normally distributed, with most seasons having between
49 and 112 wins (blue lines below). This means that there aren't too many extreme 
seasons where the team won significantly more or significantly less games than 
usual. This serves as a good gut-check for our final predictions; if the 
predicted wins are too high or too low, we know something is probably off.

```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.height = 3, fig.width = 5, fig.align = 'center'}

avgWins <- mean(train_set$TARGET_WINS)
sdev <- sd(train_set$TARGET_WINS)
l1 <- mean(train_set$TARGET_WINS) - (2*sdev)
l2 <- mean(train_set$TARGET_WINS) + (2*sdev)

# histogram of wins
p <- ggplot(train_set, aes(x=TARGET_WINS)) + 
  geom_histogram() + 
  geom_vline(aes(xintercept= avgWins),
            color="red", linetype="dashed", size=1)+
  geom_vline(aes(xintercept=l1),
            color="blue", linetype="dashed", size=1) +
    geom_vline(aes(xintercept=l2),
            color="blue", linetype="dashed", size=1) +
  labs(title="Distribution of Wins",x="Number of wins", y = "Number of seasons")

p
```

### What's missing? 
A first look at the data shows that only about 8% of the records have 
a full set of information. The good news is that most of the missing values come 
from what we would expect because they don't happen too often; hit-by-pitch 
(92% missing!), caught stealing (34% missing), and double plays (13% missing). 
Since we have so little hit-by-pitch data, we expect that it doesn't contribute 
to overall wins and will eliminate it from a few of the models we propose. The other
two stats have less than half of the data missing, so we'll need to think of a 
clever way to impute these values. The remaining missing information is from a 
combination of stolen bases and strikeouts (by batters and pitchers). **It seems** 
**completely unreasonable** to have zero strike outs in a season, so this is something 
we'll most certainly have to impute.

```{r, include = FALSE}

pct_null <- data.frame(do.call("rbind", map(train_set %>% select(-INDEX), ~mean(is.na(.)))))
colnames(pct_null) <- c('PCT_NULL')

pct_null %>%
  arrange(desc(PCT_NULL)) %>%
  filter(PCT_NULL > 0)

```

### Do the individual stats affect winning?
Intuitively, we expect that Caught stealing, Errors, Hits allowed, Homeruns allowed, 
Strikeouts by batters, and Walks allowed would all have a **negative** impact on
the total wins. In other words, as these stats increase, we would expect that the 
team is less likely to win.

However, when we look at the data for each of these statistics with respect to the 
total number of wins, we don't see a clear relationship across the board. In fact, 
only Errors and Hits allowed seem to have a negative impact on the wins. Caught stealing
and strikeouts by batters appear to be random -- whether the stat for a particular season
is high or low doesn't affect the overall number of wins.

Even more interestingly, Homeruns and Walks allowed have the *opposite* effect; as the number
of homeruns and walks that are allowed from the opposing team increase, the number of wins 
appears to increase as well!

```{r, echo = FALSE, message = FALSE, warning = FALSE}

negSet <- train_set %>%
  select(TARGET_WINS, 
         'Errors' = TEAM_FIELDING_E, 
         'Hits allowed' = TEAM_PITCHING_H, 
         'Strikeouts by batters' = TEAM_BATTING_SO, 
         'Caught stealing' = TEAM_BASERUN_CS, 
         'Walks allowed' = TEAM_PITCHING_BB, 
         'Homeruns allowed' = TEAM_PITCHING_HR)

ggplot(data = negSet %>%
  gather(-TARGET_WINS, key = "STAT", value = "VALUE"), aes(x = VALUE, y = TARGET_WINS)) + 
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(title="Number of wins with respect to recorded stats",x="", y = "Number of Wins") +
  facet_wrap(~ STAT, scales = "free")

```

We can look at the same information for the stats that we expect to have a **positive**
effect on wins: Base hits, Doubles, Triples, Homeruns, and Walks by batters, Batters
getting hit by pitches, Stolen bases, Double Plays, and Strikeouts by pitchers.

Many of these stats *do* seem to have an effect on the number of wins. Most noteably,
as the number of base hits and walks increase, the team is more likely to win. We see
a slighter, positive relationship between homeruns, doubles, triples, and stolen bases.
This makes sense when we think about it -- these things tend to happen less often in
games than pure base hits and walks. Double plays and batters hit don't appear to have
an effect on the number of wins either. Once again, this intuitively makes sense because
they are less likely to happen in a game.

One thing to note is the number of strikeouts compared to the number of wins. We can see
that there are a few outliers -- abnormally high numbers of strikeouts in a season. This
should be taken with caution, as they don't represent a typical season's stats.

```{r, echo = FALSE, message = FALSE, warning = FALSE}

posSet <- train_set %>%
  select(TARGET_WINS, 
         'Base hits' = TEAM_BATTING_H, 
         'Doubles' = TEAM_BATTING_2B, 
         'Triples' = TEAM_BATTING_3B, 
         'Homeruns' = TEAM_BATTING_HR, 
         'Walks' = TEAM_BATTING_BB, 
         'Batters hit' = TEAM_BATTING_HBP,
         'Stolen bases' = TEAM_BASERUN_SB,
         'Double plays' = TEAM_FIELDING_DP,
         'Strikeouts' = TEAM_PITCHING_SO)

ggplot(data = posSet %>%
  gather(-TARGET_WINS, key = "STAT", value = "VALUE"), aes(x = VALUE, y = TARGET_WINS)) + 
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(title="Number of wins with respect to recorded stats",x="", y = "Number of Wins") +
  facet_wrap(~ STAT, scales = "free")

```

Some of the provided statistics are well-balanced in the sense that there are
very few (or no) extreme values. **Homeruns by batters**, **Strikeouts by** 
**batters**, and **Homeruns allowed** are examples of this. Aside from

```{r, echo = FALSE, message = FALSE, warning = FALSE}

set1 <-  train_set %>%
  select('Homeruns by batters' = TEAM_BATTING_HR, 
         'Strikeouts by batters' = TEAM_BATTING_SO,  
         'Homeruns allowed' = TEAM_PITCHING_HR) %>%
  gather("stat", "value") %>%
  filter(complete.cases(.) == TRUE)

vals <- ggplot(set1 , aes(x=stat, y=value)) + 
    geom_boxplot() +
  labs(title="",x="", y = "Value") +
  facet_wrap(~stat, scale="free")

vals

# sapply(train_set %>% select(-INDEX, -TARGET_WINS), function(x){length(boxplot.stats(x)$out)})

```

```{r, echo = FALSE, message = FALSE, warning = FALSE}

# Correlation matrix with significance levels (p-value)
library("Hmisc")
res2 <- rcorr(as.matrix(train_set %>% select(-INDEX)))

# Insignificant correlation are crossed
corrplot(res2$r, type="upper", order="hclust", 
          tl.col = "black", p.mat = res2$P, sig.level = 0.01, insig = "blank")

```