### 2020-09-10 HW1/HW1_AF.Rmd
more updates

---
title: 'DATA 621---Business Analytics and Data Mining'
subtitle: 'Fall 2020---Group 2---Homework #1'
author: Avraham Adler, Samantha Deokinanan, Amber Ferger, John Kellogg,
    Bryan Persaud, Jeff Shamp
date: "9/27/2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r loadData, include=FALSE}
library(tidyverse)
library(ggplot2)
library(corrplot)
train_set <- read.csv('data/moneyball-training-data.csv')
eval_set <- read.csv('data/moneyball-evaluation-data.csv')
```

## 1. DATA EXPLORATION

### How often does the team win?

We are given a data set containing 2,276 records, each of which includes 15 statistics
and the total number of wins a team had in a given year. On average, about 50% of games 
played are won (81 games out of 162), with the best season having 146 wins and the worst 
season having 0 wins. The data is normally distributed and most years have between 49 
and 112 wins (blue lines below). The nature of the distribution means there aren't too 
many extreme seasons where wins are significantly higher or lower than usual. This serves 
as a good gut-check for our final predictions; if the predicted wins are too high or too 
low, we know something in our model is probably off.

```{r wins, echo = FALSE, message = FALSE, warning = FALSE, fig.height = 3, fig.width = 5, fig.align = 'center'}

avgWins <- mean(train_set$TARGET_WINS)
sdev <- sd(train_set$TARGET_WINS)
l1 <- mean(train_set$TARGET_WINS) - (2*sdev)
l2 <- mean(train_set$TARGET_WINS) + (2*sdev)

# histogram of wins
p <- ggplot(train_set, aes(x=TARGET_WINS)) + 
  geom_histogram() + 
  geom_vline(aes(xintercept= avgWins),
            color="red", linetype="dashed", size=1)+
  geom_vline(aes(xintercept=l1),
            color="blue", linetype="dashed", size=1) +
    geom_vline(aes(xintercept=l2),
            color="blue", linetype="dashed", size=1) +
  labs(title="Distribution of Wins",x="Number of wins", y = "Number of seasons")

p
```

### What's missing? 
A first look at the data shows that only about 8% of the records have a full set
of information. The good news is that most of the missing values come from 
statistics that don't happen too often: hit-by-pitch (92% missing!), caught 
stealing (34% missing), and double plays (13% missing). Since we have so little 
hit-by-pitch data, we expect that it doesn't contribute much to overall wins and will 
eliminate it from a few of the models we propose. The other two stats have less 
than half of the data missing, so we'll need to think of a clever way to fill in 
these values. The remaining missing information is from a combination of stolen 
bases and strikeouts (by batters and pitchers). **It seems completely unreasonable** 
to have zero strike outs in a season, so this is something we'll most certainly have 
to impute.

```{r pctNull, include = FALSE}

pct_null <- data.frame(do.call("rbind", map(train_set %>% select(-INDEX), ~mean(is.na(.)))))
colnames(pct_null) <- c('PCT_NULL')

pct_null %>%
  arrange(desc(PCT_NULL)) %>%
  filter(PCT_NULL > 0)

```

### Do the individual stats affect winning?
Intuitively, we expect that Caught stealing, Errors, Hits allowed, Homeruns allowed, 
Strikeouts by batters, and Walks allowed would all have a **negative** impact on
the total wins. In other words, as these values increase, we expect that the team 
is less likely to win.

However, when we take a closer look at the data, these relationships aren't obvious. 
In fact, only Errors and Hits allowed seem to have a negative impact on wins. Caught 
stealing and Strikeouts by batters appear to be random; this means that whether the 
stat for a particular season is high or low doesn't affect the overall number of wins.

Even more interestingly, Homeruns allowed and Walks allowed have the *opposite* 
effect; as these stats increase, so do the number of wins!

```{r negFactors, echo = FALSE, message = FALSE, warning = FALSE}

negSet <- train_set %>%
  select(TARGET_WINS, 
         'Errors' = TEAM_FIELDING_E, 
         'Hits allowed' = TEAM_PITCHING_H, 
         'Strikeouts by batters' = TEAM_BATTING_SO, 
         'Caught stealing' = TEAM_BASERUN_CS, 
         'Walks allowed' = TEAM_PITCHING_BB, 
         'Homeruns allowed' = TEAM_PITCHING_HR)

ggplot(data = negSet %>%
  gather(-TARGET_WINS, key = "STAT", value = "VALUE"), aes(x = VALUE, y = TARGET_WINS)) + 
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(title="Number of wins with respect to recorded stats",x="", y = "Number of Wins") +
  facet_wrap(~ STAT, scales = "free")
```

We can look at the same information for the stats that we expect to have a **positive**
effect on wins: Base hits, Doubles, Triples, Homeruns, Walks, Batters getting hit by 
pitches, Stolen bases, Double Plays, and Strikeouts by pitchers.

Many of these stats *do* seem to have an effect on the number of wins, most notably,
base hits and walks. We see weaker positive relationships for homeruns, doubles, triples, 
and stolen bases. This makes sense when we think about it; these things tend to happen 
less often in games than pure base hits and walks, so they don't have as much of an 
effect on winning. Finally, double plays and batters hit don't appear to have any 
correlation with the number of wins. Once again, this intuitively makes sense because 
they are less likely to happen in a game.

One thing to note is the number of strikeouts compared to the number of wins. We can see
that there are a few outliers (abnormally high numbers of strikeouts in a season). This
should be taken with caution, as they don't represent a typical season's stats.

```{r posFactors, echo = FALSE, message = FALSE, warning = FALSE}

posSet <- train_set %>%
  select(TARGET_WINS, 
         'Base hits' = TEAM_BATTING_H, 
         'Doubles' = TEAM_BATTING_2B, 
         'Triples' = TEAM_BATTING_3B, 
         'Homeruns' = TEAM_BATTING_HR, 
         'Walks' = TEAM_BATTING_BB, 
         'Batters hit' = TEAM_BATTING_HBP,
         'Stolen bases' = TEAM_BASERUN_SB,
         'Double plays' = TEAM_FIELDING_DP,
         'Strikeouts' = TEAM_PITCHING_SO)

ggplot(data = posSet %>%
  gather(-TARGET_WINS, key = "STAT", value = "VALUE"), aes(x = VALUE, y = TARGET_WINS)) + 
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(title="Number of wins with respect to recorded stats",x="", y = "Number of Wins") +
  facet_wrap(~ STAT, scales = "free")

```

### Are some stats more skewed than others?

Some of the provided statistics are well-balanced in the sense that there are
very few (or no) extreme values. **Homeruns by batters**, **Strikeouts by** 
**batters**, and **Homeruns allowed** are examples of this.

```{r summaryInfo, include = FALSE}

summary(train_set)
```

```{r fewOutliers, echo = FALSE, message = FALSE, warning = FALSE, fig.height = 3, fig.width = 5, fig.align = 'center'}

set1 <-  train_set %>%
  select('Homeruns by batters' = TEAM_BATTING_HR, 
         'Strikeouts by batters' = TEAM_BATTING_SO,  
         'Homeruns allowed' = TEAM_PITCHING_HR) %>%
  gather("stat", "value") %>%
  filter(complete.cases(.) == TRUE)

vals <- ggplot(set1 , aes(x=stat, y=value)) + 
    geom_boxplot() +
  labs(title="",x="", y = "Value") +
  facet_wrap(~stat, scale="free")

vals

```
Some things to note about each of these statistics:

* Homeruns allowed (average = ~100/year) and Homeruns by batters (average = ~106/year) 
have a very similar mid-range distribution (50% of the data lies between ~50 and 150). 
The slight difference in average stats means that teams tend to have a higher number
of homeruns than the opposition team. 
* The only thing that stands out about strikeouts by batters (average = ~736/year) is 
how nearly perfectly normal it is. 50% of the data is between about 500 and 1000 and 
there are absolutely no outliers in the dataset! This means that there were no surprisingly
high or low seasons.

Conversely, some of the stats have a very high number of outliers, indicating that there
are some seasons with some abnormally high or low values. **Walks by batters**, **Stolen**
**bases**, **Hits allowed**, and **Errors** are examples of this.

```{r manyOutliers, echo = FALSE, message = FALSE, warning = FALSE, fig.height = 5, fig.width = 5, fig.align = 'center'}

set2 <-  train_set %>%
  select('Walks by batters' = TEAM_BATTING_BB, 
         'Stolen bases' = TEAM_BASERUN_SB,  
         'Hits allowed' = TEAM_PITCHING_H,
         'Errors' = TEAM_FIELDING_E) %>%
  gather("stat", "value") %>%
  filter(complete.cases(.) == TRUE)

vals2 <- ggplot(set2 , aes(x=stat, y=value)) + 
    geom_boxplot() +
  labs(title="",x="", y = "Value") +
  facet_wrap(~ stat, scale = 'free')

vals2

#sapply(train_set %>% select(-INDEX, -TARGET_WINS), function(x){length(boxplot.stats(x)$out)})

```
Some things to note about each of these statistics:

### Are stats correlated?
```{r, echo = FALSE, message = FALSE, warning = FALSE}

# Correlation matrix with significance levels (p-value)
library("Hmisc")
res2 <- rcorr(as.matrix(train_set %>% select(-INDEX)))

# Insignificant correlation are crossed
corrplot(res2$r, type="upper", order="hclust", 
          tl.col = "black", p.mat = res2$P, sig.level = 0.01, insig = "blank")

```