### 2020-09-16 Adler_personal/Blog1.Rmd
First Blog and stopped one which is going nowhere

---
title: "Blog Post 1: Orthogonal Regression"
subtitle: "DT 621---Fall 2020"
author: "Avraham Adler"
date: "9/15/2020"
output:
  prettydoc::html_pretty:
    theme: cayman
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# What is Orthogonal Regression
Orthogonal regression is when the error between the observations, \(y_i\) and
the regression line is not measured solely along the y-axis (vertically) but 
perpendicularly---orthogonally---to the regreression line. The "error" is split
evenly between the x-axis and the y-axis. This is a special case of
[Deming regression](https://en.wikipedia.org/wiki/Deming_regression), where the
relationship between the x and y error is a constant reflecting the ratio of the
variance of their errors, not necessarily equal to one as it is in orthogonal
regression.

# Why Use Orthogonal Regression?
Orthogonal and Deming regressions are special cases of
[errors-in-variables models](https://en.wikipedia.org/wiki/Errors-in-variables_models).
These are models where there is explicit recognition that the *independant*
variables are subject to error as well. Perhaps a tool was faulty, perhaps we
ran into Heisenberg's uncertainty principle and we **cannot** measure precisely.
Regardless, these kinds of models help correct for the attenuation bias that
results when independent variables have measurement error.

# Example
Let's assume that the true relationship between \(X\) and \(Y\) is:
\[
y_i = 3.1 + 6x_i + \epsilon_i\\
\epsilon \sim \mathcal{N}(0, 2)
\]

In case one, we will generate data assuming we know the \(x\)'s exactly. In case
two, we will generate white noise to add to each \(x\), representing the
measurement error. We will then compare the results of OLS with orthogonal least
squares.

```{r createData, fig.width=8}
set.seed(33)
n <- 15
x <- rnorm(n, 0, 1)
y <- 3.1 + 6 * x + rnorm(n, 0, 2)
x2 <- x + rnorm(n, 0, 2)
par(mfrow = c(1, 2))
plot(x2, y, col = 'red', xlim = c(min(x, x2), max(x, x2)),
                                  ylim = c(min(y), max(y)),
     main = "Observations With Measurment Error")
plot(x, y, xlim = c(min(x, x2), max(x, x2)), ylim = c(min(y), max(y)),
     main = "Observations w/o Measurment Error")
```

Clearly, the measurement error will inducing attenuation error in the model. We
will now regress the observed \(y\)'s on the **observed** \(x\)'s.

```{r model, fig.width=8}
set.seed(33)
library(deming)
fit1 <- lm(y ~ x2)
fit2 <- deming(y ~ x2)
Tfit <- lm(y ~ x)
summary(fit1)
fit2
plot(x2, y, col = 'red', xlim = c(min(x, x2), max(x, x2)),
                                  ylim = c(min(y), max(y)),
     main = "Observations With Measurment Error")
abline(fit1$coefficients[[1]], fit1$coefficients[[2]], col = 'red', lty = 2)
abline(fit2$coefficients[[1]], fit2$coefficients[[2]], col = 'darkgreen', lty = 2)
abline(Tfit$coefficients[[1]], Tfit$coefficients[[2]], col = 'black', lty = 2)
```

The red line is the linear regression assuming the measured \(x\)'s have no
error. The black line is the regression through the "true" value of the \(x\)'s.
The green line is the orthogonal regression line. While we were unable to fully
recover the true regression line, using orthogonal regression did address the
attenuation error, and shifted the slope more towards the actual than to 0.