### 2020-09-17 Adler_personal/Blog2.Rmd
Blog2

---
title: "Blog Post 2: Training & Testing: Linear Regression"
subtitle: "DT 621---Fall 2020"
author: "Avraham Adler"
date: "9/17/2020"
output:
  prettydoc::html_pretty:
    theme: cayman
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, message=FALSE, warning=FALSE, error=FALSE)
```

# Training & Testing Sets
When originally taught, most students perform their regression analysis on the
totality of their data. If one is solely interested in past behavior, this is
proper. However, if one is interested in predicting future behavior, it is
better practice to reserve some of the data you have for testing purposes. This
reflects the "unknown" future data which is the true intent of the model.

# Experiment
We will create synthetic data representing both past and future observations.
A quarter of the data will be the "true" future data. This will never be seen
during the modeling process. The remaining 75% reflects the "observed"
data---the full complement of what is available to the data scientist.

Two models will be built using this data. One using all of it and the second
using A 75/25 split into a training and testing set. The selected models from
each group will then be compared against the "future" data.

# Raw Data
We will use a simple linear model as our example:
\[
y_i = 3.2 + 1.6x_{1i} + 3.4x_{2i} - 0.8x_{3i} + \epsilon_i
\]

We will call the model using glm with a Gaussian link.

```{r loadPackages}
library(caret)
```
```{r createData}
set.seed(245)
x1 <- runif(400, 0, 100)
x2 <- rnorm(400, 0, 10)
x3 <- rexp(400, 1/5)
y <- 3.2 + 1.6 * x1 + 3.4 * x2 - 0.8 * x3 + rnorm(400, 0, 10)
futIDX <- sample(400, 100)
x1f <- x1[futIDX]
x1p <- x1[-futIDX]
x2f <- x2[futIDX]
x2p <- x2[-futIDX]
x3f <- x3[futIDX]
x3p <- x3[-futIDX]
yf <- y[futIDX]
yp <- y[-futIDX]
testIDX <- sample(300, 300 * 0.25)
x1trn <- x1p[-testIDX]
x1tst <- x1p[testIDX]
x2trn <- x2p[-testIDX]
x2tst <- x2p[testIDX]
x3trn <- x3p[-testIDX]
x3tst <- x3p[testIDX]
ytrn <- yp[-testIDX]
ytst <- yp[testIDX]
```

# Linear Regression Model
This is almost unfair with regards to the training/testing split. Normally, the
validation set is used to fine tune the hyperparameters. In simple linear
regression there are none. If the second model returns better results, it is
very strong evidence that there is value in simply fitting to less data to
prevent overfitting!

## Training
```{r linearFit}
trc <- trainControl(method = 'cv', number = 10)
m1 <- train(data.frame(x1 = x1p, x2 = x2p, x3 = x3p), yp, method = 'glm',
            trControl = trc, family = gaussian)
m2 <- train(data.frame(x1 = x1trn, x2 = x2trn, x3 = x3trn), ytrn, method = 'glm',
            trControl = trc, family = gaussian)
summary(m1)
summary(m2)
```

Both models recover the parameters decently, although the intercept is somewhat
off. The last is almost always due to process risk, the natural vagaries of
stichastic simulation.

## Testing
```{r linearTest}
m1p <- predict(m1, data.frame(x1 = x1f, x2 = x2f, x3 = x3f))
m2p <- predict(m2, data.frame(x1 = x1f, x2 = x2f, x3 = x3f))
defaultSummary(data.frame(pred = m1p, obs = yf))
defaultSummary(data.frame(pred = m2p, obs = yf))
```

The second model, fit on less data, has better accuracy on the "future" data.
This can only be due to overfitting. In a future blog post, we may investigate
results when there **are** hyperparameters which can be tuned.