### 2020-09-21 HW1/HW1_Group2.Rmd
Copyediting

---
title: 'DATA 621---Business Analytics and Data Mining'
subtitle: 'Fall 2020---Group 2---Homework #1'
author: Avraham Adler, Samantha Deokinanan, Amber Ferger, John Kellogg,
    Bryan Persaud, Jeff Shamp
date: "9/27/2020"
output:
  pdf_document:
    toc: TRUE
    toc_depth: 2
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, error=FALSE, warning=FALSE, message=FALSE)
```

```{r loadData, include=FALSE}
# Libraries and variables 
library(tidyverse)
library(ggplot2)
library(corrplot)
library(tidymodels)
library(mice)
library(psych)
library(Hmisc)
library(MASS)
library(dplyr)
library(caret)

urlRemote  = "https://raw.githubusercontent.com/"
pathGithub = "aadler/DT621_Fall2020_Group2/master/HW1/data/"
fileTrain   = "moneyball-training-data.csv"
train_set = read.csv(paste0(urlRemote, pathGithub, fileTrain))

fileEval   = "moneyball-evaluation-data.csv"
eval_set = paste0(urlRemote, pathGithub, fileEval) %>% read.csv()

set.seed(9450)
```

# DATA EXPLORATION
## How Often Does The Team Win?
The data set provided comprises 2,276 records of baseball teams seasons. There
are 15 features, statistics corresponding to a team's performance in a season,
and a target variable of the team's number of wins in that season. A baseball
season contains 162 games. On average, about 50% of games played are won---81
games out of 162---with the best individual season having 146 wins and the worst
season having 0 wins. The data is near normally distributed and most years have
between 49 and 112 wins, represented by the blue lines in the histogram below.
The nature of the distribution indicates that there aren't too many extreme
seasons where wins are significantly higher or lower than usual. This serves as
a good gut-check for our final predictions; if the predicted wins are too high
or too low, we know something in our model is probably off.

```{r wins, fig.height = 2.5, fig.width = 5, fig.align = 'center'}
# DATA EXPLORATION
avgWins <- mean(train_set$TARGET_WINS)
sdev <- sd(train_set$TARGET_WINS)
l1 <- mean(train_set$TARGET_WINS) - (2 * sdev)
l2 <- mean(train_set$TARGET_WINS) + (2 * sdev)
# histogram of wins
p <- ggplot(train_set, aes(x=TARGET_WINS)) +
  geom_histogram() + 
  geom_vline(aes(xintercept = avgWins),
            color = "red", linetype = "dashed", size = 1)+
  geom_vline(aes(xintercept = l1),
            color="blue", linetype = "dashed", size = 1) +
    geom_vline(aes(xintercept = l2),
            color = "blue", linetype = "dashed", size = 1) +
  labs(title = "Distribution of Wins", x = "Number of wins",
       y = "Number of seasons")
p
```

## What's Missing? 
A first look at the data shows that only about 8% of the records have a full set
of information. The good news is that most of the missing values come from
statistics that don't happen too often: hit-by-pitch (`TEAM_BATTING_HBP`, 92%
missing!), caught stealing (`TEAM_BASERUN_CS`, 34% missing), and double plays
(`TEAM_FIELDING_DP`, 13% missing). Since we have so little hit-by-pitch data, we
expect that it doesn't contribute much to overall wins and will eliminate it
from a few of the models we propose. The other two stats have less than half of
the data missing, so we'll need to think of a clever way to fill in these
values. The remaining missing information is from a combination of stolen bases
and strikeouts by pitchers and batters (`TEAM_BASERUN_SB`, `TEAM_PITCHING_SO`,
`TEAM_BATTING_SO`). It seems **completely unreasonable** to have zero strike
outs in a season, so this is something we'll most certainly have to impute.

```{r pctNull, fig.height = 2, fig.width = 5, fig.align = 'center'}
# percent null for each variable
pct_null <- data.frame(do.call("rbind", map(train_set %>% dplyr::select(-INDEX),
                                            ~ mean(is.na(.)))))
colnames(pct_null) <- c('PCT_NULL')
totalNulls <- pct_null %>%
  mutate(VARIABLE = rownames(.)) %>%
  arrange(desc(PCT_NULL)) %>%
  filter(PCT_NULL > 0) %>%
  dplyr::select(VARIABLE, PCT_NULL)
ggplot(totalNulls, aes(x = reorder(VARIABLE, PCT_NULL), y = PCT_NULL,
                       label = round(PCT_NULL, 2))) + 
  geom_text(vjust = 0.5, hjust = -0.05)+
  geom_bar(stat = "identity") +
  ggtitle("Variables with Missing Information") +
  xlab("Statistic") + ylab("Percent Missing") + 
  coord_flip() + expand_limits(y = 1)
```

## Do the Individual Stats Affect Winning?
**Stats with an expected negative impact:** Intuitively, we expect that Caught
Stealing, Errors, Hits Allowed, Home Runs Allowed, Strikeouts by Batters, and
Walks Allowed would all have a **negative** impact on the total wins. In other
words, as these values increase, we expect that the team is less likely to win.

```{r negFactors, fig.align = 'center'}
# Variables expected to have a negative impact on wins
negSet <- train_set %>%
  dplyr::select(TARGET_WINS, 
         'Errors' = TEAM_FIELDING_E, 
         'Hits allowed' = TEAM_PITCHING_H, 
         'Strikeouts by batters' = TEAM_BATTING_SO, 
         'Caught stealing' = TEAM_BASERUN_CS, 
         'Walks allowed' = TEAM_PITCHING_BB, 
         'Home Runs allowed' = TEAM_PITCHING_HR)
ggplot(data = negSet %>%
  gather(-TARGET_WINS, key = "STAT", value = "VALUE"),
  aes(x = VALUE, y = TARGET_WINS)) + 
  geom_point() + geom_smooth(method = "lm", se = FALSE) +
  labs(title = "Number of wins with respect to recorded stats",
       x = "", y = "Number of Wins") + facet_wrap(~ STAT, scales = "free")
```

When we take a closer look at the data, these negative relationships aren't
obvious. In fact, only **Errors** and **Hits Allowed** seem to have a negative
impact on wins. **Caught Stealing** and **Strikeouts by Batters** appear to be
random; this means that whether the stat for a particular season is high or low
doesn't affect the overall number of wins.

Even more interestingly, **Home Runs Allowed** and **Walks Allowed** have the
*opposite* effect---as these stats increase, so do the number of wins!

**Stats with an expected positive impact:** We can look at the same information
for the stats that we expect to have a **positive** effect on wins: Base Hits,
Doubles, Triples, Home Runs, Walks, Batters getting hit by pitches, Stolen
Bases, Double Plays, and Strikeouts by Pitchers.

```{r posFactors, fig.align = 'center'}
# Variables expected to have a positive impact on wins
posSet <- train_set %>%
  dplyr::select(TARGET_WINS, 
         'Base hits' = TEAM_BATTING_H, 
         'Doubles' = TEAM_BATTING_2B, 
         'Triples' = TEAM_BATTING_3B, 
         'Home Runs' = TEAM_BATTING_HR, 
         'Walks' = TEAM_BATTING_BB, 
         'Batters hit' = TEAM_BATTING_HBP,
         'Stolen bases' = TEAM_BASERUN_SB,
         'Double plays' = TEAM_FIELDING_DP,
         'Strikeouts' = TEAM_PITCHING_SO)
ggplot(data = posSet %>%
  gather(-TARGET_WINS, key = "STAT", value = "VALUE"),
  aes(x = VALUE, y = TARGET_WINS)) + geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(title = "Number of Wins with Respect to Recorded Stats", x = "",
       y = "Number of Wins") + facet_wrap(~ STAT, scales = "free")
```

Many of these stats *do* seem to have an effect on the number of wins; most
notably, **Base Hits** and **Walks**. We see weaker positive relationships for
**Home Runs**, **Doubles**, **Triples**, and **Stolen Bases**. On the surface,
this makes sense as these events tend to happen less often in games than base
hits and walks, so they don't have as much of an effect on winning. Finally,
**Double Plays** and **Batters Hit** don't appear to have any correlation with
the number of wins. Once again, this intuitively makes sense because they are
less likely to happen in a game.

One thing to note is the number of strikeouts compared to the number of wins. We
can see that there are a few outliers---abnormally high numbers of strikeouts in
a season. This should be taken with caution, as they don't represent a typical
season's stats.

## Are Some Stats More Skewed Than Others?
Before using any of the statistics in a model, we need to take a closer look at
the variation in the data. We call out-of-the-ordinary values, exceptionally
high or low values, **outliers**. We need to take these into account in our
modeling to prevent them from skewing our predictions.

```{r summaryInfo, fig.height = 3, fig.width = 5, fig.align = 'center'}
# Understanding outliers by variable
totalOutliers <- data.frame(
  sapply(train_set %>% dplyr::select(-INDEX, -TARGET_WINS, - TEAM_BATTING_HBP), 
       function(x){length(boxplot.stats(x)$out)/nrow(train_set)}))
totalOutliers$VARIABLE_NM <- rownames(totalOutliers)
colnames(totalOutliers) <- c('PCT_OUTLIERS', 'VARIABLE_NM')
ggplot(totalOutliers,
       aes(x = reorder(VARIABLE_NM, PCT_OUTLIERS), y=PCT_OUTLIERS,
           label = round(PCT_OUTLIERS, 3))) + 
  geom_text(vjust = 0.5, hjust = -0.05)+ geom_bar(stat = "identity") +
  ggtitle("Percentage of Outliers") + xlab("Statistic") +
  ylab("Percent of Data that is an Outlier") + coord_flip() +
  expand_limits(y = 0.15)
```

As we can see, some of the provided statistics are well-balanced in the sense
that there are very *few*, or no, extreme values. **Home Runs allowed**
(`TEAM_PITCHING_HR`), **Strikeouts by batters** (`TEAM_BATTING_SO`), and
**Home Runs by batters** (`TEAM_BATTING_HR`) are some examples.

```{r fewOutliers, fig.height = 2, fig.width = 6, fig.align = 'center'}
# Distributions of variables with few outliers
set1 <- train_set %>%
  dplyr::select('Home Runs by batters' = TEAM_BATTING_HR, 
         'Strikeouts by batters' = TEAM_BATTING_SO,  
         'Home Runs allowed' = TEAM_PITCHING_HR) %>%
  gather("stat", "value") %>%
  filter(complete.cases(.) == TRUE)
vals <- ggplot(set1 , aes(x = stat, y = value)) + 
    geom_boxplot() + labs(title = "", x = "", y = "Value") +
  facet_wrap( ~ stat, scales = "free")
vals
```

Some things to note about each of these statistics:

  * **Home Runs Allowed** (average ~100/year) and **Home Runs by Batters**
  (average ~106/year) have a very similar mid-range distribution---50% of the
  data lies between ~50 and 150. The slight difference in average stats means
  that teams tend to have a higher number of Home Runs than the opposition team.
  * The only thing that stands out about **Strikeouts by Batters** (average
  ~736/year) is how nearly perfectly normal it is. 50% of the data is between
  about 500 and 1000 and there are absolutely no outliers in the dataset! This
  means that there were no surprisingly high or low seasons.

Conversely, some of the stats have a very *high* number of outliers, indicating
that there are some seasons with some abnormally high or low values. **Errors**
(`TEAM_FIELDING_E`), **Hits allowed** (`TEAM_PITCHING_H`), **Walks by batters**
(`TEAM_BATTING_BB`), and **Stolen bases** (`TEAM_BASERUN_SB`) are some examples.

```{r manyOutliers, fig.height = 2, fig.width = 6, fig.align = 'center'}
# Distributions of variables with many outliers
set2 <-  train_set %>%
  dplyr::select('Walks by batters' = TEAM_BATTING_BB, 
         'Stolen bases' = TEAM_BASERUN_SB,  
         'Hits allowed' = TEAM_PITCHING_H,
         'Errors' = TEAM_FIELDING_E) %>%
  gather("stat", "value") %>%
  filter(complete.cases(.) == TRUE)
vals2 <- ggplot(set2 , aes(x=stat, y=value)) + 
    geom_boxplot() +
  labs(title = "",x = "", y = "Value") +
  facet_wrap(~ stat, scales = 'free', ncol = 4)
vals2
```

Some things to note about each of these statistics:

  * All of the outliers for **Errors**, **Hits Allowed**, and **Stolen Bases**
  are above the upper tail of the data set. This is further illustrated by the
  mean and median values for these stats; in all instances, the means per year
  (Errors ~246/year, Hits Allowed ~1779/year, Stolen Bases ~125/year) are higher
  than the medians per year (Errors ~159/year, Hits Allowed ~1518/year, Stolen
  Bases ~101/year). This means that some seasons with exceptionally high values
  skew the dataset.
  * There are a few *very* extreme outliers for **Hits Allowed**. The maximum
  value is 30,132, which is over 16 times the average number of hits allowed per
  season!
  * There are outliers both above *and* below the tails of the data for the
  **Walks by Batters** stat. This means that we have exceptionally low
  (min = 0!) and exceptionally high (max = 878) seasons.

The remaining stats, **Walks Allowed** (`TEAM_PITCHING_BB`),
**Base Hits by Batters** (`TEAM_BATTING_H`), **Caught Stealing**
(`TEAM_BASERUN_CS`), **Strikeouts by Pitchers** (`TEAM_PITCHING_SO`),
**Double Plays** (`TEAM_FIELDING_DP`), **Triples by Batters**
(`TEAM_BATTING_3B`), and **Doubles by Batters** (`TEAM_BATTING_2B`) have between
29 and 99 outliers.

```{r otherOutliers, fig.height = 3.5, fig.width = 7, fig.align = 'center'}
# Distribution of all other variables 
set3 <-  train_set %>%
  dplyr::select('Base Hits by batters' = TEAM_BATTING_H, 
         'Doubles by batters' = TEAM_BATTING_2B,  
         'Triples by batters' = TEAM_BATTING_3B,
         'Caught stealing' = TEAM_BASERUN_CS,
         'Walks allowed' = TEAM_PITCHING_BB,
         'Strikeouts by pitchers' = TEAM_PITCHING_SO,
         'Double plays' = TEAM_FIELDING_DP) %>%
  gather("stat", "value") %>%
  filter(complete.cases(.) == TRUE)
vals3 <- ggplot(set3, aes(x=stat, y=value)) + 
    geom_boxplot() +
  labs(title = "", x = "",y = "Value") +
  facet_wrap(~ stat, scales = 'free', ncol = 4)
vals3
```

Some things to note:

  * All variables are very narrowly distributed, meaning that most of the data
  falls within a small range. 
  * **Strikeouts by Pitchers** and **Walks Allowed** have a few very extreme
  outliers; these   represent seasons that have abnormally high values for the
  statistics. 
  * The average number of pure **Base Hits** (1469/season) is greater than the
  average number of Doubles (~241/season) and Triples (~55/season). This isn't
  at all surprising, but serves as a good gut check on the validity of the data.

## Are Stats Correlated?
We would expect that a few things in the dataset might be correlated. Perhaps
number of errors and hits/homeruns allowed or the number of base hits by batters
and doubles/triples/homeruns. We can visualize the correlations between the
statistics to determine if there is a significant relationship between the them.
In the graph below, blue dots represent positively correlated variables and red
dots represent negatively correlated variables.

```{r corrPlot, fig.align = 'center'}
# Correlation matrix with significance levels (p-value)
res2 <- rcorr(as.matrix(train_set %>% dplyr::select(-INDEX)))
# Insignificant correlation are crossed
corrplot(res2$r, type="upper", order="hclust", 
          tl.col = "black", p.mat = res2$P, sig.level = 0.01, insig = "blank")
```

Some noteworthy relationships (coincidental or not):

  * **Errors** are highly, negatively correlated with walks by batters,
  strikeouts by batters, and homeruns (both by batters and allowed).
  * **Triples by Batters** are highly, negatively correlated with strikeouts by
  batters and homeruns (both by batters and allowed).
  * **Strikeouts by Batters** are highly, positively correlated with homeruns
  (both by batters and allowed). 
  * **Homeruns by Batters** and **Homeruns Allowed** are both positively
  correlated with walks by batters.
  * As expected, **Base Hits** are positively correlated with doubles and
  triples.

We can keep these correlations in mind when developing our models: if we have
correlated statistics, there could be in-built redundancy in the features, and
we may be able to create a simpler, more accurate model by eliminating some.

# DATA PREPARATION
Now we have a good idea of the data we are looking at we can take the next steps
to prepare it for building a solid model.

## Outlier Removal
As we saw in the data analysis, there are some outlier concerns for some of the
variables, so we will need to account for this in our modeling. When performing
the processes of outlier removal, a cautious approach is always best. Each
outlier is evaluated to ensure:

  * It is clearly from incorrectly or mis-centered data.
  * Its removal does not affect later assumptions.
  * It creates a significant association/relation that is eliminated with its removal.

We can take another look at the outliers for each variable:

```{r outliers_1, fig.height = 4, fig.width = 8, fig.align = 'center'}
# Visualizing variables prior to outlier removal
train_set %>%
  dplyr::select(-TARGET_WINS, -INDEX) %>%
  pivot_longer(everything(), 
               names_to = "Variable", 
               values_to = "Value") %>%
  ggplot(aes(x = Variable, y = Value)) +
  geom_boxplot(na.rm = TRUE) + 
  labs(title="Outliers for each Statistic",x="Variable", y = "Value") +
  coord_flip()
```

Off the bat (pun intended), there are some clear issues with two of the
variables: `TEAM_PITCHING_SO` and `TEAM_PITCHING_H.`

  * **TEAM_PITCHING_SO**: The largest outlier is close to 20,000, which averages
  ~123 strikeouts per game. This is only possible if every one of the 162 games
  went approximately 40 innings, and each out was a strikeout---which clearly
  has never happened.
  * **TEAM_PITCHING_H**: The largest outlier for this variable is over 30,000
  which is similarly highly unlikely.

Since both appear with a heavy right skew, we will use median and IQR to remove
the outliers.

```{r iqr_removal_SO}
# Remove outliers
train_set<- train_set %>%
  na.omit() %>%
  summarise(iqr = IQR(TEAM_PITCHING_SO)) %>%
  bind_cols(
            train_set 
            )  %>%
  filter(
    TEAM_PITCHING_SO > quantile(TEAM_PITCHING_SO,
                                probs = c(0.25), 
                                na.rm = TRUE) - 1.5 * iqr, 
    TEAM_PITCHING_SO < quantile(TEAM_PITCHING_SO, 
                                probs = c(0.75), 
                                na.rm = TRUE) + 1.5 * iqr
        ) %>%
  dplyr::select(-iqr)
```
```{r iqr_removal_H}
train_set<- train_set %>%
  na.omit() %>%
  summarise(iqr = IQR(TEAM_PITCHING_H)) %>%
  bind_cols(
            train_set 
            )  %>%
  filter(
    TEAM_PITCHING_H > quantile(TEAM_PITCHING_H,
                                probs = c(0.25), 
                                na.rm = TRUE) - 1.5 * iqr, 
    TEAM_PITCHING_H < quantile(TEAM_PITCHING_H, 
                                probs = c(0.75), 
                                na.rm = TRUE) + 1.5 * iqr
        ) %>%
  dplyr::select(-iqr)
```

We can take a look at the distributions of these variables after removing the
outliers:

```{r outlier_boxplot, fig.height = 4, fig.width = 8, fig.align = 'center'}
# visualizing variables post outlier removal
train_set %>%
  dplyr::select(-TARGET_WINS, -INDEX) %>%
  pivot_longer(everything(), 
               names_to ="Variable", 
               values_to="Value") %>%
  ggplot(aes(x=Variable, y=Value)) +
  geom_boxplot(na.rm = TRUE) + 
  labs(title="Outliers for each Statistic",x="Variable", y = "Value") +
  coord_flip()
```

This has eliminated what appears to be the most extreme outliers. There are
still large outlier sets for Errors and Stolen Bases, but none that seem to
dwarf the other variables. According to the box plot, observations appear to be
on a logical scale given the data.

## Imputation
Our early exploration of the data showed that the majority of the data is
complete with only a few variables with missing values. We will use the **MICE**
(Multivariate Imputation by Chained Equations) method. MICE is a principled
method for dealing with missing data. It creates multiple imputations for the
mean, as opposed to single imputations, and accounts for the statistical
uncertainty in these imputations.

```{r impute_train}
# Imputing values
train_set <- complete(mice(data = train_set, 
                         method = "pmm", 
                         seed = 9450, 
                         print=FALSE), 3)
```

## Feature Engineering
### Single base hits
With outliers removed and missing values imputed via MICE, we can create new
variables. The first will be single base hits, and it will be derived from some
of the variables we do have. We know that Team Batting Hits (`TEAM_BATTING_H`)
is a combination of *all* hits for the season, so we can create Single Base Hits
as follows:

\[
Singles = Total\ Hits - (Doubles + Triples + Homeruns)
\]

```{r singles}
# Defining singles variable 
train_set<- train_set %>%
  dplyr::select(-INDEX) %>%
  mutate_if(is.integer, as.numeric) %>%
  mutate(TEAM_BATTING_1B =
           TEAM_BATTING_H -
           (TEAM_BATTING_2B + TEAM_BATTING_3B + TEAM_BATTING_HR))
```

### Slugging Percentage
The second variable we will create is **Slugging Percentage**. Slugging is an
offensive statistic that is a good predictor of winning and it tends to have
better variance and correlation behavior than most other variables. It is
composed of singles, doubles, triples, home runs and at-bats:

\[
\mathrm{SLG} ={\frac{{\mathit{1B}} + 2\times{\mathit{2B}} +
3\times{\mathit{3B}} + 4\times {\mathit{HR}}}{\mathit{AtBats}}}
\]

Because we don't have a statistic for at-bats, we will approximate it as:

\[
\widehat{\mathrm{SLG}} = {\frac{\mathit{1B} + 2\times\mathit{2B} +
3\times\mathit{3B} + 4\times\mathit{HR}}{SO_{batting} + H + BB}}
\]

```{r make_slg_obp}
# Defining at bats and slugging variables
train_set<- train_set %>%
  mutate(AB = TEAM_BATTING_SO + TEAM_BATTING_BB + TEAM_BATTING_H,
         SLG = (TEAM_BATTING_1B + 2 * TEAM_BATTING_2B + 3 * TEAM_BATTING_3B +
                  4 * TEAM_BATTING_HR) / AB)
```

As a sanity check, we can take a look at Slugging in comparison to Triples
(`TEAM_BATTING_3B`) and Home Runs (`TEAM_BATTING_HR`). All of these variables
should have a positive correlation with the total number of wins.

```{r slugging_compare, fig.height = 2.5, fig.width = 8, fig.align = 'center'}
# Visualizing slugging compared to wins
modified_set <- train_set %>%
  dplyr::select(TARGET_WINS, SLG, TEAM_BATTING_3B, TEAM_BATTING_HR)
ggplot(data = modified_set %>%
  gather(-TARGET_WINS, key = "STAT", value = "VALUE"),
  aes(x = VALUE, y = TARGET_WINS)) + 
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(title = "Number of wins with respect to recorded stats", x = "",
       y = "Number of Wins") +
  facet_wrap(~ STAT, scales = "free")
```

# BUILD MODELS
## Model 1
### Base Model
We will start with a simple linear model to serve as a baseline. This includes
all variables in the dataset.

```{r baseModel}
# BUILD MODELS
# Model 1: Backwards Selection
lm_reg = lm(data=train_set, TARGET_WINS ~ .)

knitr::kable(summary(lm_reg)$coefficients, digits = 3L,
             caption = 'Base Model Regression Output')
```

We can immediately see that a few variables *exceed* the 0.05 p-value threshold
for significance. Additionally, we see that both `TEAM_BATTING_1B` and `AB` have
no coefficients at all! This is likely due to the fact that they are colinear
with other variables in the model.

### Enhanced Model
We used a combination of domain knowledge and data analysis to justify removing
some features. First, we are keeping the coefficient for `TEAM_BATTING_HR` since
its p-value is marginally above the general threshold and knowledge of the game
suggests it is important. We also chose to remove the `TEAM_BATTING_HBP`
variable because most of the values are missing from the dataset.

The remaining variables were chosen as a result of **backwards selection** based
on null hypothesis testing for non-zero slope. This is done by removing one
variable at a time to determine the minimum number of significant variables
needed for prediction.

Altogether, we end up **removing** 7 total variables: `TEAM_BATTING_1B`, `AB`, 
`TEAM_BATTING_3B`, `TEAM_BATTING_H`, `TEAM_PITCHING_BB`, `TEAM_PITCHING_SO`, and
`TEAM_BATTING_1B`.

### Further data transformation
We used Cook's distance to remove additional outliers that are influencing the
fit of the model above. We used a cutoff of \(\frac{4}{N}\).

```{r cooks_distance}
# removing outliers via cooks distance
lm_1<- lm(TARGET_WINS ~ TEAM_BATTING_2B + TEAM_BATTING_HR +
          TEAM_BATTING_BB + TEAM_BATTING_SO + TEAM_BASERUN_SB +
            TEAM_BASERUN_CS + TEAM_PITCHING_HR + TEAM_PITCHING_H +
            TEAM_FIELDING_E + TEAM_FIELDING_DP + SLG, data = train_set)
cooks_dis<- cooks.distance(lm_1)
influential<- as.numeric(names(cooks_dis)[(cooks_dis > (4 / nrow(train_set)))])
train_df2<- train_set[-influential, ]
lm_1<- lm(TARGET_WINS ~ TEAM_BATTING_2B + TEAM_BATTING_HR +
            TEAM_BATTING_BB + TEAM_BATTING_SO + TEAM_BASERUN_SB +
            TEAM_BASERUN_CS + TEAM_PITCHING_HR + TEAM_PITCHING_H +
            TEAM_FIELDING_E + TEAM_FIELDING_DP + SLG, data = train_df2)
mod_sum1 = summary(lm_1)  
```

```{r model1_coef}
knitr::kable(summary(lm_1)$coefficients, digits = 3L,
             caption = 'Backwards Model Regression Output')
```

### Coefficient Discussion
There are some counter intuitive results, which is expected given that baseball is a 
messy, imprecise game that has evolved over time. To that end, we should expect some 
seemingly strange results from algorithmic regression.

First, we see that the intercept itself is negative. This means that, given no information
about any of the statistics, we would predict that a team would lose (by a lot!).

Next, fielding double plays and batting doubles both appear to have negative impacts
on wins even though they *should* be positive impacts. Turning double plays,
while a good for the defensive team, may suggest a larger, negative issue.
Namely, weak pitching that leads to runners on base. Similarly, batting doubles,
leaves runners open to double plays. Allowed hits by pitching, caught stealing,
and batting strike-outs are all counter intuitive results as well, but they seem
to be small contributors and these are events that happen regularly in every
game.

Slugging as an approximation is the major predictor in this regression, by far.
The other predictors other than the intercept are orders of magnitude less in
predictive value. It should be noted that slugging alone is not a good predictor
for wins overall.

## Model 2
This is a model allowing pairwise interactions within the data types of
baserunning, batting, pitching, and fielding. It is fit using a forward and
backwards stepwise regression based on AIC.

### Further data transformation
All the features will be centered and scaled. The target variable will be left
in normal space. This will make prediction easier.

```{r model2C+S}
##### Model 2: Pairwise Model
# cleaning data for model 2
ts2 <- train_set[, -(17:19)]
ts2 <- data.frame(apply(ts2, 2, scale))
ts2$TARGET_WINS <- train_set$TARGET_WINS
```

### Stepwise Model
A fully specified model with all the interactions will be the starting point for
the stepwise regression, which uses AIC as its target metric. For interest, this
model will be compared with both the null model and the fully-specified model
from where the stepwise process begins

```{r model2}
nullModel <- lm(TARGET_WINS ~ 1, data = ts2)
fullModel <- lm(TARGET_WINS ~
  (TEAM_BASERUN_CS + TEAM_BASERUN_SB) ^ 2 + (TEAM_BATTING_2B + TEAM_BATTING_3B +
  TEAM_BATTING_BB + TEAM_BATTING_H + TEAM_BATTING_HR + TEAM_BATTING_SO) ^ 2 +
  (TEAM_FIELDING_DP + TEAM_FIELDING_E) ^ 2 + (TEAM_PITCHING_BB +
  TEAM_PITCHING_H + TEAM_PITCHING_HR + TEAM_PITCHING_SO) ^ 2, data = ts2)
stepModel <- stepAIC(fullModel, scope = list(upper = ~
  (TEAM_BASERUN_CS + TEAM_BASERUN_SB) ^ 2 + (TEAM_BATTING_2B + TEAM_BATTING_3B +
  TEAM_BATTING_BB + TEAM_BATTING_H + TEAM_BATTING_HR + TEAM_BATTING_SO) ^ 2 +
  (TEAM_FIELDING_DP + TEAM_FIELDING_E) ^ 2 + (TEAM_PITCHING_BB +
  TEAM_PITCHING_H + TEAM_PITCHING_HR + TEAM_PITCHING_SO) ^ 2, lower = ~ 1),
  direction = 'both', trace = 0)
snm <- summary(nullModel)
sfm <- summary(fullModel)
sstm <- summary(stepModel)
stmr2 <- sstm$r.squared
stmar2 <- sstm$adj.r.squared
stmAIC <- AIC(stepModel)
stmRMSE <- sqrt(mean(stepModel$residuals ^ 2))
```

### Discussion
#### Models
```{r model2_discuss}
knitr::kable(snm$coefficients, digits = 3L,
             caption = 'Null Model')
knitr::kable(sfm$coefficients, digits = 3L,
             caption = 'Full Model with All Features & Pairwise Class Interactions')
knitr::kable(sstm$coefficients, digits = 3L,
             caption = 'Stepwise Model with Pairwise Class Interactions')
```

The null model expects teams to lose slightly more than half their games. The
full model has a lower intercept, which indicates that the impact of all the
features explains why teams *win* slightly more than just the grand mean does.
The stepwise model is similar, having a lower intercept than the null but a
higher intercept that the full model.

Also, the full model has the higher \(R^2\) and the stepwise model has the
higher *adjusted* \(R^2\). This is expected, as there is no penalty for adding
more parameters in regular \(R^2\) measures.

The **stepwise** model will be the one discussed below.

#### Coefficients
Most of the coefficients are reasonable within the context of the game of
baseball. These two statements are axiomatic:

  * The only way to win is to have the higher score at the end of the game.
  * The only way to score is for a baserunner to cross home plate.

With those in mind, we can make the following observations about the linear
non-interactive coefficients.

 * Reasonable
   * Caught stealing removes baserunners but it isn't significant and can be
   ignored on its own.
   * Stolen bases get a runner closer to home plate; positive coefficient makes
   sense.
   * Triples get a runner very close to home plate; positive coefficient makes
   sense.
   * Walks are free baserunners; positive coefficient makes sense but it isn't
   significant and can be ignored on its own.
   * Hitting home runs directly increase the score; positive coefficient makes
   sense.
   * Striking out reduces the number of baserunners; negative coefficient makes
   sense.
   * Errors allow the other team free baserunners; negative coefficient makes
   sense.
   * Giving up walks allows the other team free baserunners; negative
   coefficient makes sense but it isn't significant and can be ignored on its
   own.
   * Giving up home runs gives the opponent scores; negative coefficient makes
   sense.
   * Getting strikeouts reduces the opponents baserunners; positive coefficient
   makes sense.
 * Curious
   * Hits increase the number of baserunners; why negative and insignificant?
   * Hitting doubles get runners on base; why negative?
   * Turning double plays reduce baserunners; why negative?
   * Giving up hits allows the other team baserunners; why positive although
   insignificant?

First, an absolutely fascinating observation from an earlier state of the model.
In the very first draft of this exercise, prior to some of the data adjustments
performed here, `TEAM_BATTING_HR` was removed due to its high correlation with
`TEAM_PITCHING_HR`. In that first draft, pitching giving up home runs was given
a *positive* coefficient. This made no sense. Allowing batting home runs to
re-enter the model, In hindsight this becomes obvious. The pitching
*was being used as an indicator for* ***batting!!*** They are almost 100%
correlated! This correlation allowed the use of pitching HRs as an indicator for
the hidden batting HRs which has a larger magnitude! Once both variables were
restored to the model, the logical coefficients surfaced. This is yet another
reason why models should not be trusted out of the box, but all model results
should be reviewed for sanity and sense!!

The curiosities above can *possibly* be resolved by looking at the interaction
terms. The pitching and batting home run and strikeout coefficients are an order
of magnitude greater than all the others. It is likely that the negative
coefficients in some of the batting stats and positive coefficients in some of
the pitching stats are being used to "temper" the effect of SOs and HRs.

A possible explanation for the negative coefficient for getting double plays, is
that double plays require at least two people on base. That means that the
opponent has a lot of base runners, which is very highly correlated with
scoring.

The behavior of doubles remains confusing. Perhaps its presence in a number of
the interaction terms means that the singleton variable needs to act as a
balance. Another theory may be that it's hiding something like a team's
propensity to strand runners on base. It would be interesting to see a breakdown
between the American and National leagues, as the latter tends to be somewhat
better at "small ball" and moving the runners along.

## Model 3
The final model is a Stepwise Regression with Repeated k-fold Cross-Validation,
and higher order polynomials variables were introduced into the full model.

A stepwise variable selection model is conducted to determine what are the
variables that can help predict the number of wins for the team. The method
allows variables to be added one at a time to the model, as long as the
F-statistic is below the specified \(\alpha\), in this case \(\alpha = 0.05\).
However, variables already in the model do not necessarily stay in. The steps
evaluate all of the variables already included in the model and remove any
variable that has an insignificant F-statistic. Only after this test ends, is
the best model found, that is when none of the variables can be excluded
and every variable included in the model is significant.

Here, the dependent variable is the continuous variable, `TARGET_WINS`, and the
independent variables are the full model to identify the most contributing
predictors. In addition, a robust method for estimating the accuracy of a model,
the k-fold cross-validation method, was performed evaluate the model performance
on different subset of the training data and then calculate the average
prediction error rate.

### Further data transformation
Once again, Cook's distance is use to decide if an individual entity is an extreme 
value or not.

```{r model3}

##### Model 3: K-fold Model
stepdata <- subset(train_set, select = -c(AB, SLG, TEAM_BATTING_HBP))
stepdata <- cbind(stepdata, sapply(stepdata[2:length(stepdata)], function(x) x^2))
names(stepdata) <- make.unique(names(stepdata), sep = "_")
model <- lm(TARGET_WINS ~ ., data = stepdata)
cooksd <- cooks.distance(model)
influential <- as.numeric(names(cooksd)[(cooksd > 4*mean(cooksd, na.rm = TRUE))])
stepdata.2 <- stepdata[-c(influential), ]
# Set up repeated k-fold cross-validation
set.seed(525)
train.control <- trainControl(method = "cv", number = 10)
# Train the model
step.model <- train(TARGET_WINS ~ ., data = stepdata.2, method = "lmStepAIC", 
                   trControl = train.control, trace = FALSE)
# Final model 
step.model <- lm(TARGET_WINS ~ TEAM_BATTING_H + TEAM_BATTING_2B + TEAM_BASERUN_SB + 
                  TEAM_BASERUN_CS + TEAM_PITCHING_H + TEAM_PITCHING_HR + TEAM_FIELDING_E + 
                  TEAM_FIELDING_DP + TEAM_BATTING_BB_1 + TEAM_BATTING_SO_1 + 
                  TEAM_BASERUN_CS_1 + TEAM_PITCHING_H_1 + TEAM_PITCHING_BB_1 + 
                  TEAM_FIELDING_E_1 + TEAM_FIELDING_DP_1 + TEAM_BATTING_1B_1, 
                data = stepdata.2)
# Summary of the model
mod_sum3 <- summary(step.model)
```

### Coefficient Discussion
```{r model3_discuss}
mod_sum3[["coefficients"]] %>% 
  knitr::kable(digits = 3L, 
               caption = 'K-fold Step Model with Higher Order Polynomials Output')
```

From studying each coefficient, this discussion begins with the variables that follow 
the logic of winning and are self-evident. Firstly, the model suggests that winning is 
in favor of the team batting hits more bases, (1B, 2B, 3B, HR). It further reveals that 
an increase in singles would lead to better chances of winning. A hit is a hit, and more 
hits than the opponent will inevitably lead to a win. Likewise, an increase in walks by 
the batting team is awarded first base, further increasing base counts and, thus the 
chances of winning. In terms of the base run, an increase in the number of stolen bases, 
and a decrease in caught steals would lead to a win for the batting team. These variables 
have a positive effect on the number of wins, and it is understandable as to why.

When interpreting the model based on the pitching team's plays, it is perceptible that 
walks allowed should be reduced so that the batting team hits does not increase. Moreover, 
preventing errors and increasing double plays in which two players are put out will also 
increase the chances of winning.

On the other hand, it is noteworthy that the model suggests some counter-intuitive play 
strategies. Firstly, when all variables are null, the intercept predicts that a team would 
win, with a target win of 129. Moreover, the model suggests that a decrease in double hits 
and an increase in strikeouts by the batting team will improve winning. These variables 
were kept because when a batter steps to the plate, the player is more likely to strike 
out than to get a hit. Trying to hit the ball out of the park will come with strikeouts 
but it will also increase the chances of hitting home runs (even 1B, 2B, 3B), and that is 
a pretty good exchange that most teams are willing to perform.

# SELECT MODELS
## Criteria
In order to select the best model to make predictions, we looked at the
following goodness of fit metrics:

 1. \(R^2\), which represents the proportion of the variance explained by the
 model
 2. Adjusted \(R^2\), which is a modified version of \(R^2\) taking parsimony
 into consideration
 3. *Root Mean Squared Error* (RMSE), which is the square root of the mean
 squared difference between the observation and the fitted value
 4. *Akaike Information Criterion* (AIC), which is an information-criteria based
 estimate of the Kullback-Leibler divergence from the current model to the
 "true" model. One of the nice results of the theory of information-criteria
 based measures is that there is included recognition of parsimony.

## Performance
```{r performance}
##################### SELECT MODELS
# Model accuracy
results <- data.frame(
  Models = c("Model #1: Backwards" , "Model #2: Pairwise", "Model #3: K-fold"),
  R.squared = c(mod_sum1[["r.squared"]], stmr2, mod_sum3[["r.squared"]]), 
  adj.R.squared = c(mod_sum1[["adj.r.squared"]], stmar2,
                    mod_sum3[["adj.r.squared"]]),
  RSME = c(sqrt(mean(lm_1$residuals ^ 2)), stmRMSE,
           sqrt(mean(step.model$residuals ^ 2))),
  AIC = c(AIC(lm_1), AIC(stepModel), AIC(step.model)))
results %>% knitr::kable(digits = 3L, 
                         caption = 'Performance Statistics & Error Measure of Models')
```

```{r errors}

# calculate predictions for each model 
mod1_preds <- data.frame(cbind(train_df2$TARGET_WINS, round(mod_sum1$residuals,0), train_df2$TARGET_WINS + round(mod_sum1$residuals,0) ))
colnames(mod1_preds) <- c('TARGET_WINS', 'RESIDUAL', 'PREDICTED')

mod2_preds <- data.frame(cbind(ts2$TARGET_WINS, round(sstm$residuals,0), ts2$TARGET_WINS + round(sstm$residuals,0) ))
colnames(mod2_preds) <- c('TARGET_WINS', 'RESIDUAL', 'PREDICTED')

mod3_preds <- data.frame(cbind(stepdata.2$TARGET_WINS, round(mod_sum3$residuals,0), stepdata.2$TARGET_WINS + round(mod_sum3$residuals,0) ))
colnames(mod3_preds) <- c('TARGET_WINS', 'RESIDUAL', 'PREDICTED')

# summary statistics for residuals
resid_stats <- data.frame(rbind(summary(mod1_preds$RESIDUAL), summary(mod2_preds$RESIDUAL), summary(mod3_preds$RESIDUAL)))

resid_stats$Models <- c('Model #1: Backwards' , 'Model #2: Pairwise', 'Model #3: K-fold')

```

It was interesting to compare the three models given that their performance
statistics, error measurements, and residual distributions were not drastically 
different from each other. **Model #2: Pairwise** was the worst performing model 
for all of our metrics, and was eliminated immediately. **Model #3: K-fold** accounts 
for nearly 45% of the variation in the dependent variable with the independent 
variables because its \(R^2\) is 0.456 and its adjusted \(R^2\) is = 0.450, both of 
which are acceptable for a good model. However, **Model #1: Backwards** has the 
smallest RSME of all the models. RMSE is a measure of how far off predictions are 
from the actual, giving extra weight to outliers through summing the squared
differences. This demonstrates a greater reduction in the randomness by
**Model #1: Backwards**. But its difference from **Model #3: K-fold** is only
0.022. It is hard to judge if that is a significant difference. We thus also
considered AIC. With AIC, it is not the magnitude of the value which is
important, but the difference between the various model AIC's. The magnitude is
a function of the number of observations. Accepted rules of thumb in the
statistical literature are that if the difference is greater than 10, there is
effectively no support for the model with the larger AIC
(Section 2.6, Burnham & Anderson 2002). As a result, **Model #1: Backwards**
is our selected model.

## Discussion
This is an interesting model in that given no other information, a team would be
expected to win negative games, which is clearly impossible. That being said, it
is an indication of the predictive power of the engineereed **slugging** feature
that allows small increments in slugging to result in significant shifts in
target wins. This engineered feature contains within it numerous interactions.
It combines scaled versions of some features and normalizes them for the
combination of other features. Perhaps that is why it is so significant.

## Predictions
**Model #1: Backwards** was used to predict the target wins base on the
evaluation set, once processed similarly to the training data. As the number of
games won must be integral, the values were rounded to the closest integer.

```{r eval_set, eval=FALSE}
# predict target wins
eval_set = eval_set %>% na.omit() %>%
  summarise(iqr = IQR(TEAM_PITCHING_SO)) %>%
  bind_cols(eval_set) %>%
  filter(TEAM_PITCHING_SO > quantile(TEAM_PITCHING_SO, probs = c(0.25), na.rm = TRUE) 
         - 1.5 * iqr, 
         TEAM_PITCHING_SO < quantile(TEAM_PITCHING_SO, probs = c(0.75), na.rm = TRUE) 
         + 1.5 * iqr) %>%
  dplyr::select(-iqr)

eval_set = eval_set %>% na.omit() %>%
  summarise(iqr = IQR(TEAM_PITCHING_H)) %>%
  bind_cols(eval_set) %>%
  filter(TEAM_PITCHING_H > quantile(TEAM_PITCHING_H, probs = c(0.25), na.rm = TRUE) 
         - 1.5 * iqr,
         TEAM_PITCHING_H < quantile(TEAM_PITCHING_H, probs = c(0.75), na.rm = TRUE) 
         + 1.5 * iqr) %>%
  dplyr::select(-iqr)

eval_set = complete(mice(data = eval_set, method = "pmm", seed = 9450,  print=FALSE), 3)

eval_set = eval_set %>% dplyr::select(-INDEX) %>%
  mutate_if(is.integer, as.numeric) %>%
  mutate(TEAM_BATTING_1B = TEAM_BATTING_H - (TEAM_BATTING_2B + TEAM_BATTING_3B +
                                               TEAM_BATTING_HR))

# Cleaning Test Set for Model #1
test_df1 = eval_set %>%
  mutate(AB = TEAM_BATTING_SO + TEAM_BATTING_BB + TEAM_BATTING_H,
         SLG = (TEAM_BATTING_1B + 2 * TEAM_BATTING_2B + 3 * TEAM_BATTING_3B +
                  4 * TEAM_BATTING_HR) / AB)
# Cleaning Test Set for Model #2
test_df2 = test_df1[, -(17:19)]
test_df2 = data.frame(apply(test_df2, 2, scale))
# Cleaning Test Set for Model #3
test_df3 = cbind(test_df1, sapply(test_df1[1:length(test_df1)], function(x) x^2))
names(test_df3) = make.unique(names(test_df3), sep = "_")
```

```{r predictions}

############# predict target wins

# outlier removal - TEAM_PITCHING_SO
eval_set <- eval_set %>% na.omit() %>%
  summarise(iqr = IQR(TEAM_PITCHING_SO)) %>%
  bind_cols(eval_set) %>%
  filter(TEAM_PITCHING_SO > quantile(TEAM_PITCHING_SO, probs = c(0.25), na.rm = TRUE) 
         - 1.5 * iqr, 
         TEAM_PITCHING_SO < quantile(TEAM_PITCHING_SO, probs = c(0.75), na.rm = TRUE) 
         + 1.5 * iqr) %>%
  dplyr::select(-iqr)

# outlier removal - TEAM_PITCHING_H
eval_set <- eval_set %>% na.omit() %>%
  summarise(iqr = IQR(TEAM_PITCHING_H)) %>%
  bind_cols(eval_set) %>%
  filter(TEAM_PITCHING_H > quantile(TEAM_PITCHING_H, probs = c(0.25), na.rm = TRUE) 
         - 1.5 * iqr,
         TEAM_PITCHING_H < quantile(TEAM_PITCHING_H, probs = c(0.75), na.rm = TRUE) 
         + 1.5 * iqr) %>%
  dplyr::select(-iqr)

# mice imputation
eval_set <- complete(mice(data = eval_set, method = "pmm", seed = 9450, print=FALSE), 3)

# feature engineering
eval_set <- eval_set %>% dplyr::select(-INDEX) %>%
  mutate_if(is.integer, as.numeric) %>%
  mutate(TEAM_BATTING_1B = TEAM_BATTING_H - (TEAM_BATTING_2B + TEAM_BATTING_3B +
                                               TEAM_BATTING_HR))
eval_set <- eval_set %>%
  mutate(AB = TEAM_BATTING_SO + TEAM_BATTING_BB + TEAM_BATTING_H,
         SLG = (TEAM_BATTING_1B + 2 * TEAM_BATTING_2B + 3 * TEAM_BATTING_3B +
                  4 * TEAM_BATTING_HR) / AB)

# final predictions
predictions <- round(predict(lm_1, newdata = eval_set))
summary(predictions)
predictions
```

# REFERENCES

 * Burnham, Kenneth P., and David R. Anderson. 2002. Model Selection and
 Multimodel Inference: A Practical Information-Theoretic Approach. Second.
 New York: Springer Science+Business Media, Inc.

# CODE APPENDIX
The code chunks below represent the R code called in order during the analysis.
They are reproduced in the appendix for review and comment.
```{r appendix, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, eval=FALSE)
```
```{r setup}
```
```{r loadData}
```
```{r wins}
```
```{r pctNull}
```
```{r negFactors}
```
```{r posFactors}
```
```{r summaryInfo}
```
```{r fewOutliers}
```
```{r manyOutliers}
```
```{r otherOutliers}
```
```{r corrPlot}
```
```{r outliers_1}
```
```{r iqr_removal_SO}
```
```{r iqr_removal_H}
```
```{r outlier_boxplot}
```
```{r impute_train}
```
```{r singles}
```
```{r make_slg_obp}
```
```{r slugging_compare}
```
```{r baseModel}
```
```{r cooks_distance}
```
```{r model2C+S}
```
```{r model2}
```
```{r model3}
```
```{r performance}
```
<!--```{r eval_set}
```-->
```{r predictions}
```