### 2020-09-21 Adler_personal/Blog3.Rmd
Blog Post 3

---
title: "Blog Post 3: Training & Testing: Other Models"
subtitle: "DT 621---Fall 2020"
author: "Avraham Adler"
date: "9/21/2020"
output:
  prettydoc::html_pretty:
    theme: cayman
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, message=FALSE, warning=FALSE, error=FALSE)
```

# Training and Testing with Hyperparameters
We saw in the [previous blog post](https://rpubs.com/Avraham/DT621Blog2) that
splitting data into training and testing sets in and of itself can be valuable
to prevent overfitting. In this post, we will demonstrate how it can help with
hyperparameter tuning and model selection.

We will train two sets of models. The first will train on the entire available
data set and the best performing model will be selected. The second set will
train on the "train" portion of the dataset, and the model which performs best
on the "test" portion of the dataset will be selected. The two selected models
will go head-to-head on the unobserved "future" data. The decision metric will
be RMSE.

This is not the best example as it is so simple, but it should suffice.

```{r loadPackages}
library(caret)
trc <- trainControl(method = 'cv', number = 10)
```
```{r createData}
set.seed(245)
x1 <- runif(400, 0, 100)
x2 <- rnorm(400, 0, 10)
x3 <- rexp(400, 1/5)
y <- 3.2 + 1.6 * x1 + 3.4 * x2 - 0.8 * x3 + rnorm(400, 0, 10)
futIDX <- sample(400, 100)
x1f <- x1[futIDX]
x1p <- x1[-futIDX]
x2f <- x2[futIDX]
x2p <- x2[-futIDX]
x3f <- x3[futIDX]
x3p <- x3[-futIDX]
yf <- y[futIDX]
yp <- y[-futIDX]
testIDX <- sample(300, 300 * 0.25)
x1trn <- x1p[-testIDX]
x1tst <- x1p[testIDX]
x2trn <- x2p[-testIDX]
x2tst <- x2p[testIDX]
x3trn <- x3p[-testIDX]
x3tst <- x3p[testIDX]
ytrn <- yp[-testIDX]
ytst <- yp[testIDX]
```

# Model Training
For a serious project, one-pass tuning would be insufficient. The results of
each run would be used to fine-tune the parameters.

## Random Forest
```{r rfFit}
set.seed(245)
rf1 <- train(data.frame(x1 = x1p, x2 = x2p, x3 = x3p), yp, method = 'ranger',
            trControl = trc, tuneLength = 2L)
rf2 <- train(data.frame(x1 = x1trn, x2 = x2trn, x3 = x3trn), ytrn,
            method = 'ranger', tuneLength = 2L)
plot(rf1)
plot(rf2)
```

## Support Vector Machines
```{r svmFit}
tgrid <- expand.grid(list(C = seq(0.01, 1, 0.01)))
set.seed(245)
sv1 <- train(data.frame(x1 = x1p, x2 = x2p, x3 = x3p), yp, method = 'svmLinear',
            trControl = trc, tuneGrid = tgrid)
sv2 <- train(data.frame(x1 = x1trn, x2 = x2trn, x3 = x3trn), ytrn,
             method = 'svmLinear', trControl = trc, tuneGrid = tgrid)
plot(sv1)
plot(sv2)
```

# Model Selection
## Models Trained on All the Data
```{r MS1}
min(rf1$results$RMSE)
min(sv1$results$RMSE)
```

## Using Testing Set
```{r MS2}
rf2p <- predict(rf2, data.frame(x1 = x1tst, x2 = x2tst, x3 = x3tst))
sv2p <- predict(sv2, data.frame(x1 = x1tst, x2 = x2tst, x3 = x3tst))
RMSE(rf2p, ytst)
RMSE(sv2p, ytst)
```

## Model Testing
In both cases the support vector machine outperformed the random forest. This is
almost assuredly due to the truth being a linear model. In a more complicated
case, one may see something different.

```{r Test}
sv1p <- predict(sv1, data.frame(x1 = x1f, x2 = x2f, x3 = x3f))
sv2p <- predict(sv2, data.frame(x1 = x1f, x2 = x2f, x3 = x3f))
defaultSummary(data.frame(pred = sv1p, obs = yf))
defaultSummary(data.frame(pred = sv2p, obs = yf))
```

The second model, fit on less data, has better accuracy on the "future" data.
This is also a demonstration of robsutness against overfitting.

Once again, being that there was a simple linear model underlying truth, the SVM
dominated. A benfit of using train/testing sets allows models of very different
structure to be compared across more difficult problems. Splitting the data into
two sets allows for the holdout set to act as the unknown future.