### 2020-09-25 HW3/ensemble_model_shamp.Rmd
updated model, remove old

---
title: "shamp_ensemble_model"
author: "Jeff Shamp"
date: "9/25/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      message = FALSE, 
                      warning = FALSE)
```

```{r, message=FALSE}
library(tidyverse)
library(tidymodels)
library(stacks)
library(kernlab)
library(vip)
set.seed(9450)
```

```{r}
crime_df<- read.csv("https://raw.githubusercontent.com/aadler/DT621_Fall2020_Group2/master/HW3/data/crime-training-data_modified.csv")
crime_df<- as.tbl(crime_df)
```

```{r}
crime_df<-
  crime_df %>% 
  mutate(target = as.factor(target)) %>% 
  mutate_if(is.integer, as.numeric)

data_split <- initial_split(crime_df, 
                            strata = target, 
                            prop = 0.8)
train_df<- training(data_split)
test_df<- testing(data_split)
```

```{r}
cv_folds<- vfold_cv(train_df,
                    v = 5, repeats = 1)

crime_recipe<- 
  recipe(target ~., data=train_df)

crime_wf<- 
  workflow() %>%
  add_recipe(crime_recipe)

crtl_grid<- control_stack_grid()
```

```{r}
logit_specs<- 
  logistic_reg(
    penalty = tune(),
    mixture = tune()
    ) %>%
  set_engine("glm") %>%
  set_mode("classification")

svm_specs<- 
  svm_rbf(
    cost = tune(),
    rbf_sigma = tune(),
    margin = tune()
  ) %>%
  set_mode("classification") %>%
  set_engine("kernlab")
```

```{r}
svm_wf<-
  crime_wf %>%
  add_model(svm_specs)

svm_results<-
  tune_grid(
    svm_wf, 
    resample = cv_folds,
    control = crtl_grid,
    grid = 10,
    # save_pred = TRUE, 
    # save_workflow = FALSE
  )

logit_wf<- 
  crime_wf %>%
  add_model(logit_specs)

logit_results<-
  tune_grid(
    logit_wf,
    resamples = cv_folds,
    control = crtl_grid,
    grid = 10,
    # save_pred = TRUE, 
    # save_workflow = FALSE
)
```

```{r}
crime_model_stack <- 
  stacks() %>%
  add_candidates(logit_results) %>%
  add_candidates(svm_results) %>%
  blend_predictions() %>%
  fit_members()

crime_model_stack
```

We see that the logistic classifier is the top weight in the blended model, with the support vector machine...supporting the classification results.

```{r}
theme_set(theme_bw())
autoplot(crime_model_stack)
```

```{r}
crime_pred <- 
  test_df %>%
  bind_cols(predict(crime_model_stack, ., type = "prob"), 
            predict(crime_model_stack, ., type = "class"))
```

Computing the ROC AUC for the model:

```{r, eval = FALSE}
roc<- yardstick::roc_auc(
  crime_pred,
  truth = target,
  contains(".pred_1")
  )

acc<- yardstick::accuracy(
  crime_pred, 
  truth = target, 
  estimate = .pred_class
)

recall<- yardstick::recall(
  crime_pred,
  truth = target, 
  estimate = .pred_class
)

precise<- yardstick::precision(
  crime_pred,
  truth = target, 
  estimate = .pred_class
)
metrics_df<-
  bind_rows(roc, acc, recall, precise)

crime_pred %>%
  conf_mat(truth = target, estimate = .pred_class)
```

```{r}
metrics_df
```

```{r}
logit_wf %>%
  last_fit(data_split) %>%
  collect_metrics()

logit_wf %>%
  fit(data = train_df) %>%
  pull_workflow_fit() %>%
  vip(geom = "col", aesthetics = list(fill='red4'))
```

```{r}
eval_df<- read.csv("https://raw.githubusercontent.com/aadler/DT621_Fall2020_Group2/master/HW3/data/crime-evaluation-data_modified.csv")
eval_df<-as.tbl(eval_df)
```

```{r}
eval_pred <- 
  eval_df %>%
  bind_cols(predict(crime_model_stack, ., type = "prob"), 
            predict(crime_model_stack, ., type = "class"))
```

```{r}
bind_cols(
eval_pred %>%
  filter(.pred_class == 0) %>%
  count(name = "negative"),
eval_pred %>%
  filter(.pred_class ==1) %>%
  count(name = "positive")
)
```