### 2020-10-07 HW3/HW 3_Bryan.Rmd
Update HW 3_Bryan.Rmd
Adding more exploring plots

---
title: "HW 3_Bryan"
author: "Bryan Persaud"
date: "10/6/2020"
output: html_document
---

```{r}
library(tidyverse)
library(kableExtra)
library(corrplot)
library(ggplot2)
library(GGally)
```

```{r}
train_crime <- read.csv("https://raw.githubusercontent.com/aadler/DT621_Fall2020_Group2/master/HW3/data/crime-training-data_modified.csv")
```

# Data Exploration

```{r}
summary(train_crime) %>%
  kable() %>%
  kable_styling()
```

```{r}
sapply(train_crime, function(x) 
  sum(is.na(x)))
```

There are no missing data in the dataset.

```{r}
correlation <- cor(train_crime)
correlation
```

```{r}
corrplot(correlation, diag = FALSE)
```

A correlation matrix and plot helps to see which variables are more related to each other than others.

```{r}
corrplot(correlation, type = "lower", order="hclust", tl.col = "black", tl.srt = 45, method = "number")

```

```{r, fig.width=14, fig.height=12}
ggpairs(as.data.frame(correlation), lower = list(continuous = "points", combo = "facethist", discrete = "facetbar", na =
    "na"))

```