### 2020-10-08 Adler_personal/Blog5.Rmd
Blog Post 5

---
title: "Blog Post 5: The Value of Cross Validation & Hyperparameter Tuning"
subtitle: "DT 621---Fall 2020"
author: "Avraham Adler"
date: "10/8/2020"
output:
  prettydoc::html_pretty:
    theme: cayman
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, message=FALSE, warning=FALSE, error=FALSE)
```

# What is Cross-Validation?
We saw in [previous blog posts](https://rpubs.com/Avraham/DT621Blog2) that
splitting data into training and testing sets helps to prevent overfitting. In
this post, we will demonstrate another method that can be used both separately
and in conjunction with holdout sets: **cross-validation**.

Cross validation is the process by which multiple models are fit to the same
data, each time holding out a different subset of that data. Unlike true holdout
methods, every data point is in at least one model being compared. However. this
does allow for a range of pseudo-holdout data which may better address
overfitting.

The most common types of cross validation are the *k-fold* and the
*leave-one-out* (LOO). The k-fold breaks the data into \(k\) partitions of
approximately the same size, and trains \(k\) models, each time leaving out a
different fold. LOOCV is k-fold taken to its extreme, where each fold is
composed of one data point. So if there are \(n\) observations there will be
\(n\) models.

The two overfitting prevention methods may be combined. One can sequester a
true holdout set for final model comparison and use cross-validation on the
training set for better overfitting behavior. One can even go so far as to split
the data into **three** sections: a true *test* set, a *validation* set on which
one can check the results of hyperparameter tuning, and the *train* set used to
fit the model including cross-validation.

# What is Hyperparamter Tuning?
Hyperparameters in machine learning usually refer to model-specific parameters
which are independent of the data. For example, the \(\lambda\) used in lasso
or ridge regression is a hyperparameter. The number of neighbors \(k\) is a
hyperparameter. Often, careful selection of these values can result in a model
with increased sensitivity to the data

# Demonstration
For this blog post we will use the `cars` dataset from the `caret` package, and
compare the results of using cross validation in the training phase. We will not
perform any EDA or preprocessing for the purpose of this post. We will train two
sets of models. We will select a "future" set of around 25% of the data to
reflect true unobserved data. We will split the remaining data into training and
testing at about a 70/30 ratio.

The first set of models will train on the entire available training set and the
best performing model on the test set will be selected. The second set will do
the same, but will include 5-fold cross validation and hyperparameter tuning as
part of the training The decision metric will be RMSE.

```{r prepareData}
library(caret)
data(cars)
set.seed(84)
futIDX <- createDataPartition(cars$Price, p = 0.25)$Resample1
futDat <- cars[futIDX, ]
curDat <- cars[-futIDX, ]
tstIDX <- createDataPartition(curDat$Price[-futIDX], p = 0.3)$Resample1
trnDat <- curDat[-tstIDX, ]
tstDat <- curDat[tstIDX, ]
```

# Model Training
For a serious project, one-pass tuning would be insufficient. The results of
each run would be used to fine-tune the parameters.

## ElasticNet
```{r enFit}
trc <- trainControl(method = 'none')
set.seed(245)
en1 <- train(Price ~ ., data = tstDat, method = 'glmnet', trControl = trc)
trc <- trainControl(method = 'cv', number = 5)
en2 <- train(Price ~ ., data = tstDat, method = 'glmnet', trControl = trc)
en1p <- predict(en1, tstDat)
en2p <- predict(en2, tstDat)
```

## Cubist
```{r cubFit}
trc <- trainControl(method = 'none')
tg <- expand.grid(committees = 1L, neighbors = 0)
set.seed(245)
cb1 <- train(Price ~ ., data = tstDat, method = 'cubist', trControl = trc,
             tuneGrid = tg)
trc <- trainControl(method = 'cv', number = 5)
set.seed(245)
cb2 <- train(Price ~ ., data = tstDat, method = 'cubist', trControl = trc)
cb1p <- predict(cb1, tstDat)
cb2p <- predict(cb2, tstDat)
```

# Model Selection
## Models Trained on All the Data
```{r MS1}
modRes <- cbind(data.frame(Model = c("EN", "EN-CV", "Cubist", "Cubist-CV")),
                rbind(defaultSummary(data.frame(obs = tstDat$Price, pred = en1p)),
                      defaultSummary(data.frame(obs = tstDat$Price, pred = en2p)),
                      defaultSummary(data.frame(obs = tstDat$Price, pred = cb1p)),
                      defaultSummary(data.frame(obs = tstDat$Price, pred = cb2p))
                      )
                )
knitr::kable(modRes, digits = 3L, caption = "Results on Training Set")
```

It is clear that cross-validation helps each time. The ability to compare
different tuning parameters has an extraordinary effect on the rules-based model
tree (`Cubist`). This last model would clearly be the one selected.

# Future Test
We will show results for all four models.
```{r MS2}
futRes <- cbind(data.frame(Model = c("EN", "EN-CV", "Cubist", "Cubist-CV")),
                rbind(defaultSummary(data.frame(obs = futDat$Price,
                                                pred = predict(en1, futDat))),
                      defaultSummary(data.frame(obs = futDat$Price,
                                                pred = predict(en2, futDat))),
                      defaultSummary(data.frame(obs = futDat$Price,
                                                pred = predict(cb1, futDat))),
                      defaultSummary(data.frame(obs = futDat$Price,
                                                pred = predict(cb2, futDat)))
                      )
                )
knitr::kable(futRes, digits = 3L, caption = "Results on 'Future' Set")
```

As expected, using both cross validation and hyperparameter tuning enhanced the
accuracy of both sets of models. It is also gratifying to see that the single
model we would have used was the most performant.

A real-world exercise would have started with significant EDA, have used more
models, and would have continued with more precise hyperparameter tuning.