### 2020-10-27 HW3/HW3_Final_AA.Rmd
remove unneded libraries, add crtl_grid

---
title: "DATA 621 - Business Analytics and Data Mining"
subtitle: 'Fall 2020 - Group 2 - Homework #3'
author: Avraham Adler, Samantha Deokinanan, Amber Ferger, John Kellogg,
    Bryan Persaud, Jeff Shamp
date: "10/27/2020"
output:
  pdf_document: default
urlcolor: purple
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, error=FALSE, warning=FALSE, message=FALSE)
```

```{r loadData, include=FALSE}
library(tidyverse)
library(tidymodels)
library(data.table)
library(vip)
library(summarytools)
library(corrplot)
library(knitr)
library(rsample) # model 1 libraries
library(caret)
library(geoR)

set.seed(9450)

urlRemote  = "https://raw.githubusercontent.com/"
pathGithub = "aadler/DT621_Fall2020_Group2/master/HW3/data/"
fileTrain = "crime-training-data_modified.csv"
fileTest = "crime-evaluation-data_modified.csv"

df <- read.csv(paste0(urlRemote, pathGithub, fileTrain))
nobs <- dim(df)[[1]]
DT <- as.data.table(df)
DT[, target := factor(target)]
eval <- read.csv(paste0(urlRemote, pathGithub, fileTest))
```

# DATA EXPLORATION
## Data Description
The training data set contains 466 records summarizing attributes of various
neighborhoods in the city of Boston. The response variable is coded such that it
is `1` when the neighborhood’s crime rate is above the median and `0` when it is
not. In all, there are 12 predictors. These include:

Predictor Variables|Description   
-|----
zn | proportion of residential land zoned for large lots (over 25000 square feet) 
indus | proportion of non-retail business acres per suburb  
chas | a dummy var. for whether the suburb borders the Charles River (1) or not (0)
nox | nitrogen oxides concentration (parts per 10 million) 
rm | average number of rooms per dwelling   
age | proportion of owner-occupied units built prior to 1940   
dis | weighted mean of distances to five Boston employment centers  
rad | index of accessibility to radial highways   
tax | full-value property-tax rate per $10,000 
ptratio | pupil-teacher ratio by town 
lstat | lower status of the population (percent) 
medv | median value of owner-occupied homes in $1000s

This data set has complete cases, thus there is no need for imputation. Based on
some common summary statistics, there are more observations where the crime rate
is below the median. It is already apparent that some of the predictors varies
depending the crime rate. For instance, there is a noticeable difference in the
means of `age`, `lstat`, `rad`, and `zn` between the crime rate groups.

```{r sumstat}
# DATA EXPLORATION 
summarystat = stby(data = df, INDICES = df$target, FUN = psych::describe)
kable(summarystat[[1]][-13,-c(1,7)], 
      caption = "Descriptive Statistics: Crime Rate > Median", 
      digit = 2L)
kable(summarystat[[2]][-13,-c(1,7)], 
      caption = "Descriptive Statistics: Crime Rate < Median", 
      digit = 2L)
```

## Data Distribution
For each predictors, we computed kernel density estimators to understand their
distribution. The following plots show how predictors are distributed between
areas where the crime rate is higher than the median (blue) and areas where the
crime rate is below the median (red). It is of interest to understand variables
that highlight large variations between the two groups.

```{r density, fig.width=8}
# density plots
DT[, IDX := .I]
DTM <- melt(DT, id.vars = c('IDX', 'target'), measure.vars = 1:12,
            variable.name = 'metric', value.name = 'value')
ggplot(DTM, aes(x = value, fill = target, color = target)) +
  geom_density(alpha = 0.4, show.legend = FALSE) +
  facet_wrap(~ metric, scales = 'free') 
```

The density plots reveal that most of the data does not have a normal
distribution. There are some are heavily right tailed variables and others which
are multi-modal. The most Gaussian of the variables appears to be the one
related to the average number of rooms per dwelling: `rm`. Another
interesting predictor is `zn`, the proportion of residential land zoned for
large lots. Nearly 73% of the observations have a value of 0 with the remaining
spread widely between 1% and 100%. This may the source of possible distortion. A
third highly-skewed variable is `chas`, the indicator as to whether the lot
borders the Charles River. Out of the `r nobs` observations,
`r sum(DT$chas == 0)` do **not** border the river.

To provide another view, this time highlighting outliers, the data was analyzed
using boxplots.

```{r boxplots, fig.width=8}
# boxplots

ggplot(DTM, aes(x = metric, y = value)) + 
  geom_boxplot(color = "blue", fill = "blue", alpha = 0.2, notch = TRUE,
               notchwidth = 0.8, outlier.colour = "red", outlier.fill = "red",
               outlier.size = 3) + 
  stat_summary(fun.y = mean, color = "red", geom = "point", shape = 16,
               size = 2) + coord_flip() +
  labs(title = "Boxplot of Predictor Variables") + scale_y_log10()

ggplot(DTM, aes(x = metric, y = value)) + 
  geom_boxplot(alpha = 0.5, outlier.colour = "red", outlier.fill = "red",
               outlier.size = 2, aes(fill = target)) + 
  facet_wrap( ~ metric, scales = "free") + 
  labs(title = "Boxplot of Predictor Variables by Crime Rate") +
  scale_y_log10()
```

It is clear from the second boxplot, similar to the densities, that the
distribution of the predictor variables is different for for two outcomes. This
suggests that a model will be able to extract signal from the data.

## Data Correlation
Below is a corrgram of the data. The shape of the ellipse reflects the strength
of the correlation and the color represents the direction. Thus, the diagonal
which represents the perfect correlation of each variable with itself is a blue
diagonal line. The shape and the color scheme reflect the guide at the bottom
of the image.

```{r correlation, fig.height=5, fig.width=5}
# correlations
corrplot(cor(df), method = 'ellipse', type = 'lower', order = 'hclust')
```

Looking at the corrgram, we see that there are many variables which are
moderately to highly correlated, $\left|\rho\right| > 0.50$. Of note is that
`nox` has a largest positive correlation with `target` and `dis` has the largest
negative correlation with `target`.

Perhaps this is a reflection of the noticeable influence that socio-economic
status has on the crime rate of an area. It is known that crime is concentrated
in disadvantaged, urban neighborhoods in the United States. The economic
segregation suggests that affluent neighborhoods may be further away in terms of
distance, and as a result, disadvantaged areas are more attractive to crime
because the probability of success is higher even if the targets are not as
profitable. Consider areas with a higher socio-economics, these areas are
correlated have a lower crime rate. Moreover, industrial areas are correlated
with higher levels of nitrogen oxide concentrations. This may be an indication
of why these areas are less density with residents of a higher status. This
trend is also seen with the `ptratio` since a higher ratio means less funding
for public institutions, which is common in areas of lower status.

Lastly, it seems that the variable `chas`, which indicate whether the suburb
borders the Charles River, has statistically insignificant correlation with
almost all of the other variables.

# DATA PREPARATION
There are no missing values so none need to be imputed. Different preparation
techniques may be used for different models.

## Models 1 & 2
The variable `chas` will be removed due to having statistically insignificant\
correlation with almost all the variables in the dataset. Next, outliers will be
tempered by capping from below at the 5th percentile and from above at the 95th
percentile.

```{r model1_2_adjust}
new_df <- subset(df, select = -chas)

# Cap all observations at their 5th and 95th percentiles
low5 <- apply(new_df, 2, quantile, prob = 0.05)
up95 <- apply(new_df, 2, quantile, prob = 0.95)
for (i in seq_along(new_df)) {
  new_df[, i] <- pmin(new_df[, i], up95[i])
  new_df[, i] <- pmax(new_df[, i], low5[i])
}
dens_ageO <- ggplot(new_df, aes(age)) + geom_density()
dens_lstatO <- ggplot(new_df, aes(lstat)) + geom_density()
dens_rmO <- ggplot(new_df, aes(rm)) + geom_density()

```

The Box-Cox transformation will be applied to the `age`, `lstat`, and `rm`
variables.

```{r m12BC}
ageBC <- boxcoxfit(new_df$age)
lstatBC <- boxcoxfit(new_df$lstat)
rmBC <- boxcoxfit(new_df$rm)
```

The transform suggests a \(\lambda\) of `r ageBC$lambda` for `age`,
`r lstatBC$lambda` for `lstat`, and `r rmBC$lambda` for `rm`

```{r m12applyBC}
new_df$age <- new_df$age ^ ageBC$lambda
new_df$lstat <- new_df$lstat ^ lstatBC$lambda
new_df$rm <- new_df$rm ^ rmBC$lambda
```

Below are the densities of the variables prior (left) and post (right) the
Box-Cox transform. The variables `lstat` and `rm` have benefited from the
transform, but the strong bimodality of `age` has not been affected
significantly.

```{r m12BC_plot}
dens_age <- ggplot(new_df, aes(age)) + geom_density()
dens_lstat <- ggplot(new_df, aes(lstat)) + geom_density()
dens_rm <- ggplot(new_df, aes(rm)) + geom_density()
grid.arrange(dens_ageO, dens_lstatO, dens_rmO, dens_age, dens_lstat, dens_rm,
             layout_matrix = cbind(c(0, 1, 2), c(3, 4, 5)))
```

## Model 3
For model 3, very little data manipulation will be done. There may be valuable
information in the outliers and a reasonable model may be able to extract
something from `chas`. However, `chas` will be converted to a factor. Similarly,
`rad` is an index of accessibility, so it is almost certainly a factor, although
probably ordered. Given there are now two factors in the predictors, dummy
dummy variables need to be created. Lastly, the **labels** of `target` will be
changed from integer to character for compliance with some of R's naming
conventions. Based on the definitions, 0 will be mapped to *below* and 1 to
*above*.

# BUILD MODELS
First, we will split the training data into two sets. This will allow us to
perform a cross validation scheme on the models to tune parameters for optimum
performance.

```{r m12_tts}
# BUILD MODELS
# train-test split

crime_df <- new_df %>%
  mutate(target = as.factor(target)) %>% 
  mutate_if(is.integer, as.numeric)

# This is so Model 3 can use the same train/test split as Models 1 and 2
crime_df$id <- seq_len(nrow(crime_df))

# Changing the text above caused changes in output as fewer command were run. It
# is ALWAYS a good idea to set a seed IMMEDIATELY before any sampling in a
# Markdown file to ensure repeatability.

set.seed(7104)
data_split <- initial_split(crime_df, 
                            strata = target, 
                            prop = 0.8)
train_df <- training(data_split)
test_df <- testing(data_split)

# Capture the indices and remove the no-longer-necessary column
tstIDX <- test_df$id
crime_df <- subset(crime_df, select = -id)
train_df <- subset(train_df, select = -id)
test_df <- subset(test_df, select = -id)
```

## Model 1
### Base Model
We will start with a simple logistic regression model to serve as a baseline.
This includes all variables in the dataset.

```{r baseModel1}
baseModel <- glm(target~., family = binomial, data = train_df)
kable(summary(baseModel)$coefficients, digits = 3L,
      caption = 'Base Model Logistic Regression Output')
```

We can immediately see that a few variables *exceed* the 0.05 p-value threshold
for significance.

### Enhanced Model
We will use **backwards stepwise regression** to remove features that are not
statistically significant in predicting the target. The result is a model that
includes the following features: `zn`, `nox`, `age`, `dis`, `rad`, `tax`,
`ptratio`, `lstat`, and `medv`.

### Coefficient Discussion
It is important to note that the coefficients from the model are predicting
whether or not the target variable is a **1** (the crime rate is *above* the
median value). Additionally, it must be noted that the numeric coefficients are
relative to the range of values that the variable encompasses. What this means
is that it's possible to have a coefficient that seems small when we look at the
absolute magnitude, but that actually has a very strong effect when applied to
the data.

```{r model1_coeff}
# 10-fold cross validation
train_control <- trainControl(method = "cv", number = 10)

# train the model on training set
model1 <- train(target ~ zn + nox + age + dis + rad + tax + ptratio + lstat + medv,
               data = train_df,
               trControl = train_control,
               method = "glm",
               family=binomial())
kable(summary(model1)$coefficients, digits = 3L,
      caption = 'Backwards Model Regression Output')
```

Let's take a look at the *signs* of the coefficients. We can see that the
coefficients for the variables `zn` and `tax` are negative. This is indicative
of an inverse relationship; i.e., the higher these values, the less likely the
crime rate is above the median. The relationship for the zone variable aligns
with the findings from our initial data exploration. However, the relationship
for the tax variable does not; in fact, we saw that there was a positive
correlation between the two (as the tax rate increases, so does the probability
that the crime rate will be above the median).

The rest of the variables have positive coefficients. Many of these are
expected. For example, in our exploratory data analysis, we noticed that the
`nox` variable (nitrogen oxides concentration (parts per 10 million)) has the
greatest positive correlation with the target variable. This is surprising; we
hadn't thought that the nitrogen oxide concentration would have as large of an
impact on the crime rate, but some quick googling shows us that there may
actually be a relationship
([https://www.sciencedaily.com/releases/2019/10/191003114007.htm](Source)).

## Model 2
### Logistic Classifier

```{r m2setup}
cv_folds <- vfold_cv(train_df, v = 10, repeats = 1)

crime_recipe <- recipe(target ~., data=train_df)

crime_wf<- workflow() %>%
  add_recipe(crime_recipe)
```

```{r m2svm}
logit_specs <-
  logistic_reg(
    penalty = tune(),
    mixture = tune()
    ) %>%
  set_engine("glm") %>%
  set_mode("classification")

```

Tuning the models for best parameters.

```{r m2stack}
logit_wf <-
  crime_wf %>%
  add_model(logit_specs)

crtl_grid<-control_grid(verbose = FALSE)

logit_results <-
  tune_grid(
    logit_wf,
    resamples = cv_folds,
    control = crtl_grid,
    grid = 10,
    save_pred = TRUE,
    save_workflow = FALSE
)
```

```{r m2predict}
best_model_auc <- select_best(logit_results, "roc_auc")
final_auc<-
  finalize_workflow(
    logit_wf,
    best_model_auc
)

final_model_pred<-
  final_auc %>%
  last_fit(data_split) %>%
  collect_predictions()

```

Computing the metrics for the logit model:

```{r m2Stats}
roc<- yardstick::roc_auc(
  final_model_pred,
  truth = target,
  contains(".pred_1")
  )

acc<- yardstick::accuracy(
  final_model_pred, 
  truth = target, 
  estimate = .pred_class
)

recall<- yardstick::recall(
  final_model_pred,
  truth = target,
  estimate = .pred_class
)

precise<- yardstick::precision(
  final_model_pred,
  truth = target,
  estimate = .pred_class
)
metrics_df<-
  bind_rows(roc, acc, recall, precise)

crime_pred %>%
  conf_mat(truth = target, estimate = .pred_class)
```

## Model 3
```{r m3trainTemp}
DT_train <- DT[!(IDX %in% tstIDX)]
```

Model 3 will be a binary logistic regression allowing for all variables and
selected pair-wise interactions pruned using both forward and backward stepwise
regression based on AICc to select the optimal parameters. The captured indicies
of the test set for models 1 & 2 will be used to split the data for model 3 as
well. To address the possibility of overfitting, five-fold cross-validation will
be done on the training set. With the vanilla logistic regression model in R,
`glm` with a binomial family and a logit link, there are no other
hyperparameters available for tuning. More sophisticated logistic regression
implementation have these parameters for fine-tuning.

To consider *all* interactions would be foolhardy. Given the need for dummy
variables, there will 19 predictors and so \(_{19}C_2 = 171\) possible pairs.
This is too much. Therefore, intentional selection of pairs is required.
Looking at the predictors, there are some which logically may interact with
each other, such as:

  * `indus` and `nox`
    * Nitrous Oxide is a common industrial pollutant. There is a high
    correlation in the training set between the two. Allowing for its
    interaction may temper the result of their relationship without losing more
    information

What may prove interesting is to look at some unsupervised clustering techniques
to provide us with insights as to which variables may be "close" enough to
warrant interactions. Instead of treating the observations as samples, we will
treat the predictors as samples!

```{r m3Clust, fig.height=6}
predClust <- hclust(dist(t(DT_train[, -c('target', 'IDX')]), diag = TRUE))
plot(predClust, xlab = "Raw Predictors")
```

It seems that there may be reason to consider an interaction between `chas` and
`nox`. Can it be that many industrial plants are on the river? Similarly with
`indus` and `lstat`. Higher-cost homes tend to be further away from the traffic
and pollution of industrial areas.

Our selected starting model will include the following interactions:

  * indus:nox
  * indus:lstat
  * chas:nox
  * ptratio:medv
  * rm:dis

### Train Model

```{r m3dataPrep}
DT[, `:=`(chas = factor(chas, labels = c('OffRiver', 'OnRiver')),
          rad = factor(rad),
          target = factor(target, labels = c('Below', 'Above')))]
DT_test <- DT[IDX %in% tstIDX]
DT_train <- DT[!(IDX %in% tstIDX)]
```
```{r m3build}
trC <- trainControl(method = 'cv', number = 5, classProbs = TRUE,
                    summaryFunction = twoClassSummary)
dVars <- dummyVars(target ~ indus * nox +
                     indus * lstat +
                     chas * nox +
                     ptratio * medv +
                     rm * dis + . - IDX, data = DT_train, fullRank = TRUE)
DTtrnx <- predict(dVars, newdata = DT_train)
DTtrny <- DT_train$target
set.seed(1)
m3Fit <- train(x = DTtrnx, y = DTtrny, method = 'glmStepAIC', trControl = trC,
               family = 'binomial', trace = 0, direction = 'both')
```

### Coefficient Discussion
```{r m3Sum}
kable(summary(m3Fit$finalModel)$coefficients, digits = 3L,
             caption = 'Stepwise Interaction-Allowed Logistic Regression Output')
```

The selected model has some interesting results. Firstly, the predictor with the
greatest effect is `nox`. Further, the only model variable which serves to
**decrease** the crime rate from the baseline is `rm`. Other variables, such as
`chas`

The interactions between `indus` and both `nox` and `lstat` are significant. As
theorized, these interactions serve to temper their individual contributions to
crime. Remember, a 1 indicates crime above the median, so positive coefficients
indicate contributions to *increasing* the crime rate. A similar observation
holds `rm`. While `dis` on its own did not decrease the model's AICc, its
interaction to temper `rm` did help the model's deviance. No other selected
interactions were left in the model.

A number of variables fell out of the model, such as `chas`, `dis`, and `zn`
amobg others. The remaining predictors are all significant, at least at the 10%
level, except for `rad.24`. However, from the last trace of the stepping routine
shown below, removing it would ***drastically*** reduce the models performance.
Lastly, while `rad.1` is the base case, `rad.2` and `rad.6` have disappeared as
well. This indicates that both `rad = 2` and `rad = 6` should be considered
indistinguishable from `rad = 1`. 
```
                    Df Deviance    AIC
<none>                   90.565 122.56
+ dis                1   89.191 123.19
+ zn                 1   89.558 123.56
+ tax                1   89.700 123.70
- ptratio            1   93.763 123.76
+ age                1   90.156 124.16
+ `nox:chasOnRiver`  1   90.224 124.22
- `rm:dis`           1   94.254 124.25
+ chas.OnRiver       1   90.273 124.27
+ rad.6              1   90.315 124.31
- lstat              1   94.364 124.36
+ `ptratio:medv`     1   90.494 124.49
+ rad.2              1   90.554 124.55
- rm                 1   95.001 125.00
- `indus:lstat`      1   95.633 125.63
- `indus:nox`        1   97.351 127.35
- indus              1   97.597 127.60
- medv               1   98.568 128.57
- rad.3              1  101.594 131.59
- rad.5              1  105.594 135.59
- rad.7              1  118.832 148.83
- nox                1  134.786 164.79
- rad.8              1  149.967 179.97
- rad.4              1  152.757 182.76
- rad.24             1  152.797 182.80
```

# SELECT MODELS
## Criteria
In binary classification, often, the measures of focus are precision and recall,
as opposed to accuracy. A measure combining this is the F1 score, defined as
twice the sum of precision and recall divided by their product. Another oft-used
metric is the area under the receiver operating curve, AUC. We will look at
the three metrics of accuracy, F1, and AUC, and select the model that performs
best in at least two of the three metrics.

## Performance 
```{r comparison}
compTable <- data.frame(Models = c('Model 1', 'Model 2', 'Model 3'),
                        ACC = double(3),
                        F1 = double(3),
                        AUC = double(3))
```

### Model 1
```{r model1_perf}
# Model 1 Performance
model1_preds <- predict(model1, test_df, type = "raw")
model1_probs <- predict(model1, test_df, type = "prob")
colnames(model1_probs) <- c('pred0', 'pred1')

model1_results <- test_df %>%
  bind_cols(pred = model1_preds, model1_probs)

# Metrics
m1_roc <- yardstick::roc_auc(
  model1_results,
  truth = target,
  pred0 # select the prob class that corresponds to first level of target
)

m1_acc <- yardstick::accuracy(
  model1_results, 
  truth = target, 
  estimate = pred
)

m1_recall <- yardstick::recall(
  model1_results,
  truth = target, 
  estimate = pred
)

m1_precise <- yardstick::precision(
  model1_results,
  truth = target, 
  estimate = pred
)

m1_metrics_df<-
  bind_rows(m1_roc, m1_acc, m1_recall, m1_precise)
m1_metrics_df
m1CM <- confusionMatrix(model1_preds, model1_results$target,
                        mode = 'prec_recall')
compTable[1, 2] <- m1_acc$.estimate
compTable[1, 3] <- 2 * m1_precise$.estimate * m1_recall$.estimate /
  (m1_precise$.estimate + m1_recall$.estimate)
compTable[1, 4] <- m1_roc$.estimate
m1CM$table
```

### Model 2

Below the feature importances for the logistic model.

```{r}
logit_wf %>%
  last_fit(data_split) %>%
  collect_metrics()

logit_wf %>%
  fit(data = train_df) %>%
  pull_workflow_fit() %>%
  vip(geom = "col", aesthetics = list(fill='red4'))
```

We again see that the primary predictor in the logistic regression is the `nox`
with `rad` and `ptratio` rounding out the top three predictors. This is in line
with the base model. Living close to the center of town, having teachers, and
not having polluted air are the largest predictors in terms of above median crime.

### Model 3

```{r m3TestSet}
tstdVars <- dummyVars(target ~ indus * nox +
                        indus * lstat +
                        chas * nox +
                        ptratio * medv +
                        rm * dis + . - IDX, data = DT_test, fullRank = TRUE)
DTtstx <- predict(tstdVars, newdata = DT_test)
DTtsty <- DT_test$target
m3Pred <- predict(m3Fit, newdata = DTtstx)
m3PredP <- predict(m3Fit, newdata = DTtstx, type = 'prob')
m3CM <- confusionMatrix(m3Pred, DTtsty, mode = 'prec_recall')
m3CM$table
compTable[3, 2] <- sum(diag(m3CM$table)) / sum(m3CM$table)
compTable[3, 3] <- m3CM$byClass[7]
compTable[3, 4] <- yardstick::roc_auc_vec(DTtsty, m3PredP[, 1])
kable(compTable, digits = 3L, caption = 'Model Test Results')
```

# APPENDIX
The code chunks below represent the R code called in order during the analysis. They are reproduced in the appendix for review and comment.
```{r appendix, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, eval=FALSE)
```
```{r loadData}
```
```{r sumstat}
```
```{r density}
```
```{r boxplot}
```
```{r correlation}
```
```{r model1_2_adjust}
```
```{r m12BC}
```
```{r m12applyBC}
```
```{r m12BC_plot}
```
```{r model1_coeff}
```
```{r m3trainTemp}
```
```{r m3Clust}
```
```{r m3dataPrep}
```
```{r m3Sum}
```
```{r comparison}
```
```{r model1_perf}
```
```{r m3TestSet}
```