### 2020-11-19 Final_project/Draft_EDA_start.Rmd

---
title: "Final Project"
author: "John Kellogg"
date: "11/17/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(tidymodels)
library(data.table)
library(corrplot)
library(vip)
library(psych)
library(summarytools)
library(readr)
library(knitr)
library(GGally)
```

## Research question

If we know the number of likes, can we predict the number of views a video will receive?

sub questions:
For each x increase/decrease in views, dislikes, and comments, how much does the view count move?
Does turning off comments and ratings affect and by how much
Is that number affected by country?
Is this the same across categories?

## Data Exploration

```{r cars}
CA_path = "https://raw.githubusercontent.com/aadler/DT621_Fall2020_Group2/master/Final_project/Data/CAvideos.csv"
CA_set <- read_csv(file = CA_path)
US_path = "https://raw.githubusercontent.com/aadler/DT621_Fall2020_Group2/master/Final_project/Data/USvideos.csv"
US_set <- read_csv(file = US_path)
GB_path = "https://raw.githubusercontent.com/aadler/DT621_Fall2020_Group2/master/Final_project/Data/GBvideos.csv"
GB_set <- read_csv(file = GB_path)
FR_path = "https://raw.githubusercontent.com/aadler/DT621_Fall2020_Group2/master/Final_project/Data/FRvideos.csv"
FR_set <- read_csv(file = FR_path)
```

Canada data

```{r summary_ca}
summaryca <- CA_set%>% select(views, category_id,likes, dislikes, comment_count, comments_disabled, ratings_disabled)
describe(summaryca)

ggpairs(
  summaryca,
  lower = list(continuous = ggally_points, combo = ggally_dot_no_facet)
  )

ca_corr <- cor(summaryca)
corrplot(ca_corr, order = "hclust", tl.col = "black", tl.srt = 45, method = "ellipse", bg="black")

```

```{r summary_us}
summaryus <- US_set%>% select(views, category_id,likes, dislikes, comment_count, comments_disabled, ratings_disabled)
describe(summaryus)

ggpairs(
  summaryus,
  lower = list(continuous = ggally_points, combo = ggally_dot_no_facet)
  )
us_corr <- cor(summaryus)
corrplot(us_corr, order = "hclust", tl.col = "black", tl.srt = 45, method = "ellipse", bg="black")

```

```{r summary_us}
summarygb <- GB_set%>% select(views, category_id,likes, dislikes, comment_count, comments_disabled, ratings_disabled)
describe(summarygb)

ggpairs(
  summarygb,
  lower = list(continuous = ggally_points, combo = ggally_dot_no_facet)
  )

gb_corr <- cor(summarygb)
corrplot(gb_corr, order = "hclust", tl.col = "black", tl.srt = 45, method = "ellipse", bg="black")
```

```{r summary_fr}
summaryfr <- FR_set%>% select(views, category_id,likes, dislikes, comment_count, comments_disabled, ratings_disabled)
describe(summaryfr)

ggpairs(
  summaryfr,
  lower = list(continuous = ggally_points, combo = ggally_dot_no_facet)
  )

fr_corr <- cor(summaryfr)
corrplot(fr_corr, order = "hclust", tl.col = "black", tl.srt = 45, method = "ellipse", bg="black")
```

There is definite heteroskedasticity across all data points, we will need to address.