### 2020-11-26 HW5/HW5-Draft.Rmd
Minor copyediting and specifying amsmath LaTeX package

---
title: "DATA 621 - Business Analytics and Data Mining"
subtitle: 'Fall 2020 - Group 2 - Homework #5'
author: Avraham Adler, Samantha Deokinanan, Amber Ferger, John Kellogg,
    Bryan Persaud, Jeff Shamp
date: "11/28/2020"
output:
  pdf_document:
    toc: TRUE
    toc_depth: 4
    extra_dependencies:
      amsmath: null
      inputenc: utf8
urlcolor: purple
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, error=FALSE, warning=FALSE, message=FALSE)
```

```{r loadData, include=FALSE}
# Load necessary libraries
library(ggplot2)
library(knitr)
library(MASS)
library(caret)
library(corrplot)
library(mice)
library(pscl)
library(mpath)
library(data.table)

# Set master seed
set.seed(65408)

# Set filepaths for data ingestion
urlRemote  = "https://raw.githubusercontent.com/"
pathGithub = "aadler/DT621_Fall2020_Group2/master/HW5/data/"
fileTrain = "wine-training-data.csv"
fileEval = "wine-evaluation-data.csv"

# Read training file
DT <- fread(paste0(urlRemote, pathGithub, fileTrain))

# Number of training observations
ntrnobs <- dim(DT)[[1]]

# Get the names of the predictor variables
nmtrn <- names(DT)[-(1:2)]
nmtrnINT <- c('AcidIndex', 'LabelAppeal', 'STARS')
nmtrnDUB <- setdiff(nmtrn, nmtrnINT)
```

# ASSIGNMENT
The assignment for HW5 is to analyze and model a dataset containing
approximately 12,000 records representing commercially available wines. The
`TARGET` response variable represents the number of sample cases of wine
purchased by wine distribution companies after sampling that wine. These cases
would be used to provide tasting samples to restaurants and wine stores around
the United States. The more sample cases purchased, the more likely is a wine to
be sold at a high end restaurant. A large wine manufacturer is studying the data
in order to predict the number of wine cases ordered based upon the wine
characteristics. If the wine manufacturer can predict the number of cases, then
that manufacturer will be able to adjust their wine offering to maximize sales.

The objective of this assignment is to build a count regression model to predict
the number of cases of wine that will be sold given certain properties of the
wine. *HINT:* Sometimes, the fact that a variable is missing is actually
predictive of the target. You can only use the variables given to you (or
variables that you derive from the variables provided).

# DATA EXPLORATION
## Variables
The data is composed of the following variables:

|VARIABLE NAME|DEFINITION|THEORETICAL EFFECT|
|--|----|---|
|INDEX|Identification Variable (do not use)|None|
|TARGET|Number of Cases Purchased|None|
|AcidIndex|Proprietary method of testing total acidity of wine by using a weighted average||
|Alcohol|Alcohol Content||
|Chlorides|Chloride content of wine||
|CitricAcid|Citric Acid Content||
|Density|Density of Wine||
|FixedAcidity|Fixed Acidity of Wine||
|FreeSulfurDioxide|Sulfur Dioxide content of wine||
|LabelAppeal|Marketing Score indicating the appeal of label design for consumers. High numbers suggest customers like the label design. Negative numbers suggest customers don't like the design.|Many consumers purchase based on the visual appeal of the wine label design. Higher numbers suggest better sales.|
|ResidualSugar|Residual Sugar of wine||
|STARS|Wine rating by a team of experts. 4 Stars = Excellent, 1 Star = Poor|A high number of stars suggests high sales|
|Sulphates|Sulfate content of wine||
|TotalSulfurDioxide|Total Sulfur Dioxide of Wine||
|VolatileAcidity|Volatile Acid content of wine||
|pH|pH of wine||

There are `r ntrnobs` observations. All of these predictors are numeric, 
although `LabelAppeal`, `AcidIndex` and `STARS` all appear to be ordinal factors
and not true numerics. For the purpose of this assignment, we can treat them
as integers.

## Missing Data
There are a lot of missing values for some of the predictors.

```{r missingVal}
missingPreds <- transpose(DT[, lapply(.SD, function(x) {sum(is.na(x))}),
                             .SDcols = nmtrn],
                          keep.names = 'Predictors')
setnames(missingPreds, 'V1', 'Missing')
missingPreds[, Percentage := Missing / ntrnobs * 100]
setorder(missingPreds, -Missing)
kable(missingPreds, digits = 2L, caption = 'Predictors with Missing Observations')
```

`STARS` is especially sparse, but as noted in the assignment, this may be an
indication in and of its own. It is reasonable to assume that the vintners of a
good wine want it to be rated. If they do not submit it for rating, that may be
an indication of their lack of faith in the wine. How to factor this into the
analysis will be decided individually by the modelers in this assignment. The
other missing variables are all less than 10%, and their handling via imputation
or otherwise will be done on a model-by-model basis as well.

## Summary Statistics and Graphs
As all of the data are numeric, we can investigate the distributions of the
predictors both tabularly and graphically.

The numeric predictor variables have the following summary statistics, ignoring
missing values:

```{r statsN}
# Isolate numeric only predictors
predictorDT <- DT[, .SD, .SDcols = nmtrn]

# Melt them from wide to long format
predictorDTM <- melt(predictorDT, variable.name = 'metric',
                     value.name = 'value',  variable.factor = FALSE,
                     measure.vars = nmtrn)

# Calculate summary statistics
statsN <- predictorDTM[, .(Mean = mean(value, na.rm = TRUE),
                           SD = sd(value, na.rm = TRUE),
                           Min = min(value, na.rm = TRUE),
                           Q1 = quantile(value, prob = 0.25, na.rm = TRUE),
                           Median = median(value, na.rm = TRUE),
                           Q3 = quantile(value, prob = 0.75, na.rm = TRUE),
                           Max = max(value, na.rm = TRUE),
                           IQR = IQR(value, na.rm = TRUE)), keyby = 'metric']

# Print the table
kable(statsN, digits = 3L, align = 'r',
      caption = "Summary Statistitics for Numeric Variables")
```

A kernel-smoothed density plot of the distributions of the non-integer numeric
predictors is below, followed by a histogram of the integral-valued predictors.

```{r graphsD, fig.height=5, fig.width=7}
# Using Epanechnikov kernel to generate kernel-smoothed densities
ggplot(predictorDTM[!is.na(metric) & metric %chin% nmtrnDUB], aes(x = value)) +
  geom_density(kernel = 'epanechnikov') +
  facet_wrap( ~ metric, scales = 'free') +
  ggtitle('Kernel-smoothed Density of Numeric Predictors')
```

The numeric predictors all look to be basically symmetrical, but non-Gaussian in
that there is a strong spike near the median, but the tails on either side are
thicker than would be for a normal distribution with that low of a standard
deviation. This is also why a boxplot would be a poor graphical indicator, as
the spike at the medians would lead to a compressed inter-quartile range and a
plethora of outliers.

```{r graphsH, fig.height=2, fig.width=7}
# Freedman-Diaconis rule for bin widths
FDbin <- function(x) {
  result <- 2 * IQR(x, na.rm = TRUE) / (length(x) ^ (1 / 3))
  return(ifelse(result == 0, 0.5, result))
}
# Generate histogram
ggplot(predictorDTM[!is.na(metric) & metric %chin% nmtrnINT], aes(x = value)) +
  geom_histogram(binwidth = FDbin, fill = 'indianred4') +
  facet_wrap( ~ metric, scales = 'free') +
  ggtitle('Histogram of Integral Predictors')
```

Of these three, `LabelAppeal` seems to be the most "normal" of the lot; `STARS`
and `AcidIndex` look to be more Poisson in shape.

## Correlations
The corrgram below graphically represents the correlations between the numeric
predictor variables, when ignoring the missing variables.

```{r corrgram, fig.width=5, fig.height=6.5}
# Create corrgram
corrplot::corrplot(cor(DT[, ..nmtrn], use = 'complete.obs'),
         method = 'ellipse', type = 'lower', order = 'hclust',
         hclust.method = 'ward.D2')
```

There is very little correlation between the variables. The only pairs with some
level correlation are:

  * `STARS` being positively correlated with `LabelAppeal`
    * This is interesting. Could it be that wine connoisseurs are impacted by
    the visual appearance of the label and not just the flavor?
  * `AcidIndex` having some positive correlation with `FixedAcidity`
    * One may have suspected a higher correlation, to be frank.
  * `AcidIndex` having some *negative* correlation with `STARS`
    * Wine reviewers may not like too much acidity, it seems.

## TARGET variable
Lastly, the complete dataset exhibits an average of
`r prettyNum(mean(DT$TARGET), digits = 3L)` cases being bought, with the
distribution below:

```{r targetH, fig.height=2, fig.width=3, fig.align='center'}
# Plot distribution of TARGET
ggplot(DT[, .(TARGET)], aes(x = TARGET)) +
  geom_histogram(binwidth = FDbin, fill = 'darkolivegreen4')
```

# DATA PREPARATION
First, preparation necessary for all the models equally will be performed. The
subsequent imputation and feature generation, if necessary will be discussed in
a separate subsection for each modeler.

## Training & Testing Split
All the models will be trained on the same approximately 70% of the training
set, reserving 30% for validation of which model to select for the count
estimation on the supplied evaluation set.

```{r trainTestSplit}
# Create training and testing split
set.seed(1004)
trnIDX <- createDataPartition(DT$TARGET, p = 0.7)
trnSet <- DT[trnIDX$Resample1, ]
tstSet <- DT[!trnIDX$Resample1, ]
```

## Poisson Model #1 & Negative Binomial Model #1
### Missing Data
The missing data can be split into two groups: `STARS` and the rest. For all the
others, I do not think there is signal encoded in the "missingness" and thus
will use some form of imputation. `STARS` is different as discussed in the DATA
EXPLORATION section. Therefore, I will create a new feature called `Rated` which
will be true if `STARS` is missing, and then only use this variable and its
interaction with `STARS` in the model, and not `STARS` by itself.

```{r model1addVars}
# P1 & NB1: Copy the training set to leave a pristine version for others
m1trn <- copy(trnSet)

# P1 & NB1: Add the "rated" factor variable
m1trn[, Rated := factor(ifelse(is.na(STARS), 'Unrated', 'Rated'),
                        levels = c('Unrated', 'Rated'),
                        labels = c('Unrated', 'Rated'))]
```

### Imputation
Imputation will be handled through bagging. Instead of looking at a nearest
neighbors approach, which also requires centering and scaling, once can create a
set of bagged trees. As per Kuhn (2019):

>*For each predictor in the data, a bagged tree is created using all of the*
*other predictors in the training set. When a new sample has a missing*
*predictor value, the bagged model is used to predict the value. While, in*
*theory, this is a more powerful method of imputing, the computational costs*
*are much higher than the nearest neighbor technique.*

The bagged trees will use `STARS` as a predictor, as that contains valuable
information, but prior to the imputation, the missing `STARS` values will be
replaced by 0 so as not to be imputed, as per the explanation above. The
preprocessing will also check for near-zero value predictors.

```{r model1impute}
# P1 & NB1: Create bagged-tree imputation model
set.seed(89)
m1trnI <- preProcess(m1trn, method =c('nzv', 'bagImpute'))

# P1 & NB1: Replace missing STARS with 0s
m1trn[is.na(STARS), STARS := 0L]

# P1 & NB1: Impute the remaining NAs
m1trnImp <- predict(m1trnI, newdata = m1trn)
```

It will be instructive to compare the shapes of the distributions after
imputation to those before imputation.

```{r m1postImputeD, fig.height=5, fig.width=7}
# P1 & NB1: Isolate numeric only predictors
m1predictorDT <- m1trnImp[, .SD, .SDcols = nmtrn]

# P1 & NB1: Melt them from wide to long format
m1predictorDTM <- melt(m1predictorDT, variable.name = 'metric',
                     value.name = 'value',  variable.factor = FALSE,
                     measure.vars = nmtrn)

# P1 & NB1: Using Epanechnikov kernel to generate kernel-smoothed densities
ggplot(predictorDTM[metric %chin% nmtrnDUB], aes(x = value)) +
  geom_density(kernel = 'epanechnikov') +
  facet_wrap( ~ metric, scales = 'free') +
  ggtitle('Kernel-smoothed Density of Numeric Predictors')
```

Thankfully, none of the shapes appears to have changed significantly, which
implies the imputation was in line with the original distributions. Now for the
histograms.

```{r m1postImputeH, fig.height=2, fig.width=7}
# P1 & NB1: Generate histogram
ggplot(m1predictorDTM[metric %chin% nmtrnINT], aes(x = value)) +
  geom_histogram(binwidth = FDbin, fill = 'indianred4') +
  facet_wrap( ~ metric, scales = 'free') +
  ggtitle('Histogram of Integral Predictors')
```

Here too, `AcidIndex` and `LabelAppeal` look very similar to their
pre-imputation distribution. What stands out is `STARS`. Both the Poisson and
the negative binomial models are either unimodal or have to consecutive values
sharing the mode, due to their inherent nature as discrete distributions. What
is seen here is that 1 star is empirically less common than 0 (unrated) or 2
stars. On the one hand, this could simply be noise (process variance) or
parameter error (finite sample size). On the other, it could indicate the need
for a zero-inflated version of the counting distributions.

## Poisson Model #2 & Multiple Linear Regression Model #1

The right balance of the chemical properties of the wine is essential to make it
taste, look, and smell delightful, in addition to having a long shelf-life.
Typically, citric acid and alcohol are the main positive attributes a wine
should have, while volatile acidity and sulfates are the main negative ones. 
However, the dataset has negative values instead of near-zero in some of these 
variables describing the chemical properties of the wine. Because little
information is given whether or not these variables were normalized, absolute
values will be taken assuming there was an error made during data collection.

Lastly, to handle missing data for `STARS`, they will be replaced by 0, while 
the chemical properties will be imputed. Because the predictive mean matching 
imputation method combines the standard linear regression and the
nearest-neighbor imputation approach, Model Set #2 elected to fill in the
missing values of a continuous variable by using mean-matching imputation.
Afterward, the data is pre-processed to fulfill the assumption of normality
using the Yeo-Johnson transformation (2000). This technique attempts to find the
value of lambda that minimizes the Kullback-Leibler distance between the normal
distribution and the transformed distribution. This method has the advantage of
working without having to worry about the domain of \(x\).

```{r m2clean}
# P2 & ML 1: Clean and impute
set.seed(525)
m2train <- copy(trnSet[,-1])
m2train[,c(2:8,11:12)] <- abs(m2train[,c(2:8,11:12)])
m2train$STARS[is.na(m2train$STARS)] <- 0
m2train.impute <- mice(m2train, method = 'pmm', print = FALSE)
m2train <- complete(m2train.impute)
m2train.r <- m2train$TARGET
m2train.p <- m2train[,-1]
pre.process <- preProcess(m2train.p, method = c("YeoJohnson"))
m2train.p <- predict(pre.process, m2train.p)
m2train[,c(2:13,15)] <- m2train.p[,c(1:12,14)]
```

## Multiple Linear Regression #2 & Negative Binomial Model #2
### Missing Data
These models will deal with missing numerical data by imputing the NA's
using the Multivariate imputation by chained equations (MICE) method. Multiple
imputation involves creating multiple predictions for each missing value, helping 
to account for the uncertainty in the individual imputations.

```{r model3impute}
# ML2 & NB2 data transformation
# MICE imputation on train and test set
set.seed(123)
m3train <- complete(mice(data = trnSet[, -c('INDEX')],
                         method = "pmm", print = FALSE), 3)
m3test <- complete(mice(data = tstSet[, -c('INDEX')],
                         method = "pmm", print = FALSE), 3)
```

# BUILD MODELS
## Poisson Model #1 & Negative Binomial Model #1
The same person is building these two models, and the model setups will be
similar. Each model will follow a similar format, with only the family changing.
Each will start with the saturated model including all the individual predictors
***except*** for `STARS`. There will also be the following interactions:

  * `Rated` and `STARS`
    * As discussed above, by having `Rating` on its own and this interaction
    the effect of both having a rating and the effect of the rating level can be
    measured, without penalizing unrated wines a second time for a 0.
  * The full four-way interaction between `AcidIndex`, `FixedAcidity`,
  `VolatileAcidity`, and `pH`
    * There are **four** separate predictors relating to acidity. Despite their
    Pearson product-moment correlations being low, I am suspicious of too much
    weight being given to acidity. By including their interactions, effects can
    be tempered, or magnified, as necessary. `CitricAcid` is not included here
    as is not merely an acid but a flavor provider as well. Also, five-way
    interactions lead to migraines.
  * `FreeSulfurDioxide` and `TotalSulfurDioxide`
    * Again, a case of multiple theoretically related predictors.
  * `LabelAppeal` with `Rated` and `LabelAppeal` with `Rated:STARS`
    * This is a later addition. The top three predictors in magnitude were
    `LabelAppeal`, `Rated`, and `Rated:STARS`. With all three being so strong,
    there may be interactions between them. However, as `STARS` alone has been
    removed from contention, only the interactions between the two others will
    be considered.

### Poisson Model # 1
For the Poisson model, a forward and backwards stepwise procedure based on AIC
will be used, and the model with the lowest AIC on the training set will be the
selected to be tested against the testing set. Using stepAIC precludes the use
of k-fold cross validation.

```{r m1PTrain}
# P1: Using stepAIC so turn off cross-validation
trc <- trainControl(method = 'none')
set.seed(350047)
m1PFit <- train(TARGET ~ . + FreeSulfurDioxide:TotalSulfurDioxide +
                  AcidIndex * FixedAcidity * VolatileAcidity * pH +
                  Rated:STARS + LabelAppeal:Rated + LabelAppeal:Rated:STARS -
                  INDEX - STARS, data = m1trnImp,
                trControl = trc, method = 'glmStepAIC',
                family = poisson(link = 'log'), direction = 'both', trace = 0)
# P1: Save AIC
m1PFitAIC <- AIC(m1PFit$finalModel)
```

\small
```{r m1PTable}
# P1:Print results. This needed to be in its own section so that the LaTeX
# commands to change the text size can be wrapped around it.
kable(summary(m1PFit$finalModel)$coefficients,
      caption = "Model 1 Poisson Regression Output")
```
\normalsize

#### Model Checking

```{r m1PFitCheck}
# P1: Check the dispersion of the model
m1PFitDisp <- sum(residuals(m1PFit$finalModel, type = 'pearson') ^ 2) /
  m1PFit$finalModel$df.residual
```

A Poisson model should have a dispersion parameter close to 1, as the Poisson by
definition has its mean and variance equal. This model has a dispersion
parameter of `r prettyNum(m1PFitDisp, digits = 3L)` which is deemed acceptable.

#### Coefficient Discussion
Similar to the master data set, the training data set has an empirical mean
purchased cases of `r prettyNum(mean(m1trnImp$TARGET), digits = 3L)`. Having no
other information, the Poisson model expects about
`r prettyNum(exp(m1PFit$finalModel$coefficients[[1]]), digits = 3L)` cases, due
to the intercept value of `r m1PFit$finalModel$coefficients[[1]]`. This means
that the model identifies factors which lean more to *increasing* purchases than
to decreasing purchases.

The absolute magnitude of the coefficients makes sense. Factors related to
acidity tend to lower the number of cases bought. The factors with the greatest
magnitude effect are all positive, and they are the fact of being rated, and the
interactions between that and the rating level and label appeal. Keeping
everything else equal, merely being rated increases the number of cases by about
\(e^{0.6677926} \approx  1.95\)! Each additional star is predicted to contribute
another 1.2 cases to the total purchase. Interestingly, `LabelAppeal` has fallen
out of the model; its contributions are through its interactions. As surmised,
looking at all three individually may have proven too strong, as the three-way
interaction coefficient tempers the others.

All the coefficients are significant at at least the 5% level except for
Sulphates, but in terms of AIC, leaving it in provided a better model than did
taking it out.

#### Variable Importance

```{r m1PvarImp}
# P1: Extract the Poisson variable importance and order it for display
vIP <- varImp(m1PFit$finalModel)
vIPn <- row.names(vIP)
vIPDT <- data.table(Coefficients = vIPn, vIP)
setorder(vIPDT, -Overall)
kable(vIPDT, digits = 2L, caption = "Poisson Model Variable Importance")
```

The variable importance can also be seen from the trace of the step AIC
procedure. The further down the list below a variable preceded by a `-` is, the
more important it is, as its removal would cause a greater disruption to the
deviance and the AIC. The three rating variables, being rated, the rating level,
and the label appeal, are the most important variables when it comes to future
case purchases.

It is also interesting to note that there are a number of variables whose
inclusion as predictors would result in a lower deviance model than the
selected, but said drop in deviance was not enough to offset the possible
parameter error. These include `VolatileAcidity`, `LabelAppeal`, `pH`,
`Density`, `ResidualSugar`, `VolatileAcidity:pH`, and `CitricAcid`.

```
                                              Df Deviance   AIC
<none>                                             9550.7 31902
+ VolatileAcidity                              1   9549.4 31903
- Sulphates                                    1   9553.5 31903
- `FreeSulfurDioxide:TotalSulfurDioxide`       1   9553.7 31903
+ LabelAppeal                                  1   9549.9 31903
+ pH                                           1   9550.3 31904
+ Density                                      1   9550.4 31904
+ ResidualSugar                                1   9550.4 31904
+ `VolatileAcidity:pH`                         1   9550.5 31904
+ CitricAcid                                   1   9550.6 31904
- FixedAcidity                                 1   9554.6 31904
+ `FixedAcidity:pH:AcidIndex`                  1   9550.7 31904
+ `FixedAcidity:VolatileAcidity`               1   9550.7 31904
+ `FixedAcidity:pH`                            1   9550.7 31904
+ `FixedAcidity:VolatileAcidity:pH`            1   9550.7 31904
+ `VolatileAcidity:pH:AcidIndex`               1   9550.7 31904
+ `FixedAcidity:VolatileAcidity:AcidIndex`     1   9550.7 31904
+ `FixedAcidity:VolatileAcidity:pH:AcidIndex`  1   9550.7 31904
- `FixedAcidity:AcidIndex`                     1   9554.9 31904
- Alcohol                                      1   9555.1 31904
- Chlorides                                    1   9555.2 31904
- `pH:AcidIndex`                               1   9555.4 31905
- FreeSulfurDioxide                            1   9558.6 31908
- `LabelAppeal:STARS:RatedRated`               1   9565.2 31914
- TotalSulfurDioxide                           1   9568.7 31918
- `VolatileAcidity:AcidIndex`                  1   9569.6 31919
- AcidIndex                                    1   9585.3 31935
- `LabelAppeal:RatedRated`                     1   9719.7 32069
- `STARS:RatedRated`                           1  10152.1 32501
- RatedRated                                   1  10247.3 32597
```

### Negative Binomial Model # 1
As the dispersion from the Poisson model was less than 1, at first blush, it is
not expected that the negative binomial model will outperform the Poisson.

```{r m1NBTrain}
# NB1: There is no built-in AIC stepping in caret, so using basic glm.nb and
# then calling stepAIC on saturated model. Once done, investigated some one-off
# possible enhancements due to issues with glm.nb
# See https://stackoverflow.com/questions/11749977/why-does-glm-nb-throw-a-missing-value-error-only-on-very-specific-inputs
# For speed, only final called model is evaluated.
set.seed(872)
m1NBFitStart <- glm.nb(TARGET ~ . + FreeSulfurDioxide:TotalSulfurDioxide +
                         AcidIndex * FixedAcidity * VolatileAcidity * pH +
                         Rated:STARS + LabelAppeal:Rated +
                         LabelAppeal:Rated:STARS - INDEX - STARS,
                       data = m1trnImp, link = log)
m1NBFitStep <- stepAIC(m1NBFitStart, direction = 'both', trace = 0)

# Last call from automated procedure
# TARGET ~ FixedAcidity + VolatileAcidity + Chlorides + FreeSulfurDioxide + 
#     TotalSulfurDioxide + pH + Sulphates + Alcohol + LabelAppeal + 
#     AcidIndex + Rated + FreeSulfurDioxide:TotalSulfurDioxide + 
#     FixedAcidity:AcidIndex + VolatileAcidity:AcidIndex + Rated:STARS + 
#     LabelAppeal:Rated + LabelAppeal:Rated:STARS

# Removing VolatileAcidity improved the AIC

m1NBFit <- glm.nb(formula = TARGET ~ FixedAcidity + Chlorides +
                    FreeSulfurDioxide + TotalSulfurDioxide + pH + Sulphates +
                    Alcohol + LabelAppeal + AcidIndex + Rated +
                    FreeSulfurDioxide:TotalSulfurDioxide +
                    FixedAcidity:AcidIndex + VolatileAcidity:AcidIndex +
                    Rated:STARS + LabelAppeal:Rated + LabelAppeal:Rated:STARS,
                  data = m1trnImp, link = log)

# NB1: Save AIC
m1NBFitAIC <- AIC(m1NBFit)
```

\small
```{r m1NBTable}
# NB1: Print results. This needed to be in its own section so that the LaTeX
# commands to change the text size can be wrapped around it.
kable(summary(m1NBFit)$coefficients,
      caption = "Model 1 Negative Binomial Regression Output")
```
\normalsize

#### Coefficient Discussion
Similar to the Poisson model, having no other information, the negative binomial
model expects about `r prettyNum(exp(m1NBFit$coefficients[[1]]), digits = 3L)`
cases, due to the intercept value of `r m1NBFit$coefficients[[1]]`. This means
that this model as well identifies factors which lean more to *increasing*
purchases than to decreasing purchases.

The absolute magnitude of the coefficients makes less sense in this model than
in the Poisson. The factors related to appeal do increase the number of cases
bought. However, the factors related to acidity tend to show more back-and-forth
interactions than in the Poisson model.

Also, there are more "insignificant" coefficients in this model, but these are
all variables for which the AIC would increase should they have been removed.

#### Variable Importance

```{r m1NBvarImp}
# NB1: Extract the NegBinom variable importance and order it for display
vINB <- varImp(m1NBFit)
vINBn <- row.names(vINB)
vINBDT <- data.table(Coefficients = vINBn, vINB)
setorder(vINBDT, -Overall)
kable(vINBDT, digits = 2L,
      caption = "Negative Binomial Model Variable Importance")
```

The trace of the step procedure is less complete when using the `glm.nb`
function in `R` and is not listed here.

Lastly, the deviance for the negative binomial model is actually lower than that
for the Poisson model, but its AIC is higher. Based on AIC only, the Poisson
model's parsimony beats out the negative binomial's explanatory powers.

```{r m1PNB}
# P1 & NB1: GoF Comparison
m1PNBC <- data.table(Model = c("Poisson 1", "Negative Binomial 1"),
                     Deviance = c(m1PFit$finalModel$deviance,
                                  m1NBFit$deviance),
                     AIC = c(m1PFit$finalModel$aic,
                             m1NBFit$aic))
kable(m1PNBC, digits = 3L,
      caption = "Poisson 1 and NegBinom 1 Goodness-of-Fit Comparison")
```

## Poisson Model #2 & Multiple Linear Regression Model #1
The `TARGET` variable has 2734 observations with zero values. This could be a
result of the manufacturer not releasing the wine into the market, or that there
were no orders placed. Therefore, backward stepwise variable elimination for a
zero-inflated Poisson (ZIP) analysis will be investigated to understand the
conditions. This model is further compared to the same Poisson model to
determine whether there is an improvement.

In the ZIP model, the independent variable, \(TARGET_i\), take zero values
\(TARGET_i \sim 0\) with the probability \(\omega_i\) or values from Poisson
distribution \(TARGET_i \sim \mathrm{Pois}(\lambda_i)\) with probability
\((1 - \omega_i)\). For \(i = 1,\ldots, n\), it can be written as:

\[
P(Y_i = y_i) = 
\begin{cases}
\omega_i + (1-\omega_i)e^{-\lambda_i})&\text{ if } y_i = 0\\
(1-\omega_i)\frac{e^{-\lambda_i}\lambda_i^{y_i}}{y_i!}&\text{ if } y_i > 0
\end{cases}
\]

Therefore, the zero-inflated Poisson model have two parameters, \(\lambda_i\)
and \(\omega_i\), and linked with the predictor variables with the following:

\[
\begin{aligned}
&\log\left(\frac{\omega_i}{1-\omega_i}\right) = \sum_{j=1}^{t}\gamma_{ji}Z_{ji}\\
&\log\left(\lambda_i\right) = \sum_{j=1}^{k}\beta_{ji}X_{ji}
\end{aligned}
\]

where \(Z_1,\ldots, Z_l\) are the dependent variables for the first equation and 
\(X_1,\ldots, X_k\) for the second one. The expected value and variance of the
number of cases purchased for the \(i^{\textrm{th}}\) wine in the ZIP model are
respectively:

\[
\begin{aligned}
\mathrm{E}(Y_i) &= \lambda_i(1 - \omega_i)\\
\mathrm{Var}(Y_i) &= (1 - \omega_i)(\lambda_i - \omega_i\lambda_i^2)
\end{aligned}
\]

Similar to Poisson regression, the ZIP model assumes that the average number of 
cases purchased equals the variance. For more information, refer to 
[Zero-Inflated Poisson Regression](https://math.usu.edu/jrstevens/biostat/projects2013/rep_ZIP.pdf).

### Poisson Model #2

```{r m2poisson}
set.seed(525)
# Zero Inflated Poisson Model
m2.pmodel <- zeroinfl(TARGET~., 
                      dist = 'poisson', 
                      model = TRUE, 
                      data = m2train,
                      link = "logit")
# Backward Elimination
m2.fpmodel <- be.zeroinfl(m2.pmodel, 
                          data = m2train, 
                          dist = 'poisson', 
                          alpha = 0.05, 
                          trace = FALSE)
m2.poisson.aic <- AIC(m2.fpmodel)
```

Zero-inflated count models are two-component mixture models combining a point 
mass at zero with a proper count distribution. Thus, there are two sources of 
zeros: zeros may come from both the point mass and from the count component.

\small
```{r m2PTable}
# P2: Outputs
kable(summary(m2.fpmodel)$coefficients$count,
      caption = "Zero Inflated Poisson Model #2 Count Output")
kable(summary(m2.fpmodel)$coefficients$zero,
      caption = "Zero Inflated Poisson Model #2 Zero Output")
```
\normalsize

#### Model Checking

```{r m2PFitCheck}
# Dispersion statistic
E2 <- resid(m2.fpmodel, type = "pearson")
N  <- nrow(m2train)
p  <- length(coef(m2.fpmodel))  
dp <- sum(E2^2) / (N - p)

# Poisson Model #2, same formula from ZIP model
m2.poisson <- glm(TARGET ~ Alcohol + LabelAppeal + AcidIndex + STARS | VolatileAcidity + 
                    CitricAcid + FreeSulfurDioxide + TotalSulfurDioxide + pH + 
                    Sulphates + Alcohol + LabelAppeal + AcidIndex + STARS, 
                  family = poisson(link = 'log'), data = m2train)
vuong(m2.poisson, m2.fpmodel)
```

This model has a dispersion parameter of `r prettyNum(dp, digits = 3L)` which 
suggests under-dispersion (i.e., variance < mean). This parameter indicates how 
many times larger the variance is than the mean. Since the dispersion was less 
than one, the conditional variance is smaller than the conditional mean. Models 
that ignore under-dispersion will be overly conservative and may fail to detect 
significant patterns. Generalized Poisson (GP) and Conway‐Maxwell‐Poisson (CMP)
distributions are better choices for such modeling because they can handle both 
over-dispersion and under-dispersion. But for the scope of this assignment, 
it's apparent, from the Vuong test with the same Poisson model and the ZIP
model, that the zero-inflated Poisson model is a slight improvement over the
Poisson model. Therefore, the ZIP model will still be considered among others.

#### Coefficient Discussion
All of the predictors in both the count and inflation portions of the model are 
statistically significant. This model fits the data significantly better than
the null model, i.e., the intercept-only model. The count model contains
information about variables that the parent Poisson model used to estimate
`TARGET` on the condition that `TARGET` > 0. The coefficients for all 4
regression variables are statistically significant at a 99% confidence level.
The coefficient for `AcidIndex` is negative, meaning that as the total acidity
of wine goes up by a unit, the number of purchased cases goes down by 2%. One
the other hand, for the other variables, i.e. `Alcohol`, `LabelAppeal` and
`STARS`, as these increase, the number of purchased cases will also increase. In
particular, `LabelAppeal` can account for a 26% increase in the number of wine
cases sold if the visual appeal of the wine scores increase by one unit and all
other variables held constant. Whereas, there is a 19% increase in cases sold
for each increase in the wine rating by a team of experts, `STARS`.

The zero model contains information about variables that the nested Logistic 
Regression model has been used to estimate the probability of whether or not
cases are purchased. Examining the variables with a negative regression
coefficient, this suggests that as the factor for the specific chemical
properties increases, the probability of no case being purchase decreases. This
is in line with our intuition, particularly for the sulfur dioxide levels. When
a wine has an unusually small percentage of the Free and Total \(SO_2\), this
suggests the wine is chemically and/or microbially unstable. Thus, if a wine
manufacturer were to increase the sulfur dioxide levels by one unit, the odds
that the wine would certainly not be purchased would decrease by a factor of
0.92 to 0.95. However, `LabelAppeal` is among the positive coefficients for the
zero model. This is interpreted as the higher a rating is for a bottle of wine,
the more likely there will be no case purchase, and it doesn't seem reasonable.

### Multiple Linear Regrssion Model #1
Multiple Linear Regression Model #1 will be using the step function has an
option to use a process that both adds and removes variables in the model. It
uses AIC (Akaike information criterion) as a selection criterion. A full model
and a null are defined, and then the function will follow a procedure to find
the model with the lowest AIC.

```{r m1MLRFit}
# Multiple Linear Regression model 
set.seed(525)
model.null = lm(TARGET ~ 1, data = m2train)
model.full = lm(TARGET ~ ., data = m2train)

# Step model used to derive final model, m1MLRFit  
# step(model.null,
#      scope = list(upper = model.full),
#      direction = "both",
#      data = m2train, 
#      trace = FALSE)

m1MLRFit = lm(TARGET ~ STARS + LabelAppeal + AcidIndex + TotalSulfurDioxide +
                VolatileAcidity + FreeSulfurDioxide + Alcohol + CitricAcid + 
                Chlorides + pH + Sulphates,
              data = m2train)
```

\small
```{r m1MLRTable}
# ML1: Outputs
kable(summary(m1MLRFit)$coefficient,
      caption = "Multiple Linear Regression Model #1 Output")
```
\normalsize

#### Coefficient Discussion
Contributing coefficients are those which lie within the 95% level of
significance. A positive coefficient indicates that as the value of the
predictor increases, the response variable also tends to increase. A negative
coefficient suggests that as the predictor increases, the response variable
tends to decrease.

The intercept itself suggests that with no other information about the wine,
there will be cases sold. The other positive predictors suggest that increasing
`STARS` and `LabelAppeal` will result in an increase in the number of case
purchased. These make sense intuitively. Moreover, as discussed in the Poisson
Model #2, the right balance of $SO_2$ lead to better wine, thus the coefficients
are rather small and have a smaller effect when it is increased. This reasoning
is similar for `Alcohol` levels. Next, `CitricAcid`, commonly used as an acid
supplement during fermentation, helps winemakers boost the acidity of their wine
and stabilize ferric haze. So, it makes sense that the higher `CitricAcid` is,
the more cases would be sold. But once again, these need to be balance right.

On the contrary, an increase in a unit of volatile acidity results in the
decrease of case purchases. Volatile acidity is caused by bacteria or certain
yeasts. This also can increase the acid index and pH if it is high. This is why
sulfur dioxide is used to carefully control it. In all, this model should not be
taken as fixed, when it comes to chemical processes, the wine manufacturer needs
to take care of what to increase and decrease in order to reach equilibrium.

## Negative Binomial Model #2 & Multiple Linear Regrssion Model #2
### Negative Binomial Model #2
#### Baseline Model 
We will start with a simple negative binomial regression model to serve as a 
baseline. This includes all variables in the dataset.

```{r m2NBTrainBase}
# NB2: Baseline Model
m2NBBase <- glm.nb(TARGET ~ ., data = m3train)
knitr::kable(summary(m2NBBase)$coefficients, digits = 3L,
             caption = 'Negative Binomial Regression Model #2 - Base Output')
```

Immediately, we can see a few variables that do not meet the 0.05 p-value
threshold.

#### Enhanced Model
Backwards stepwise regression is performed and the result is a model with the 
following variables *removed*: `FixedAcidity`, `CitricAcid`, `ResidualSugar`, 
and `Density`. Because the `Sulphates` variable is very close to the 0.05
p-value threshold of significance, we will remove it as well.

```{r m2NBTrain}
# NB2: Revised Model
m2NBFit <- glm.nb(TARGET ~ VolatileAcidity + Chlorides +
                FreeSulfurDioxide + TotalSulfurDioxide + pH + Alcohol + 
                LabelAppeal + AcidIndex + STARS, data = m3train)
knitr::kable(summary(m2NBFit)$coefficients, digits = 3L,
             caption = 'Negative Binomial Regression Model #2 - Final Output')
```

#### Coefficient Discussion
The intercept of the final model tells us that there is typically ~1 case of
wine purchased. Some things to note:

  * Stars and label appeal contribute the most positively to the number of cases
  of wine purchased. Intuitively, this makes sense - the higher the rating and
  the prettier the bottle, the more cases we would expect to be sold.
  * Free Sulfur Dioxide and Total Sulfur Dioxide have a very tiny effect on the
  number of cases sold (rounded to 3 decimal places, the impact is 0!)
  * All of the remaining variables except alcohol have a negative effect on the
  number of cases purchased.

### Multiple Linear Regrssion Model # 2
One drawback to traditional multiple linear regression models is that the 
forecast counts are not integers. We will deal with this by rounding the
predictions to the nearest whole number.

#### Baseline Model
We will start with a simple linear model to serve as a baseline. This includes
all variables in the dataset.

```{r m2MLTrainBase}
# ML2: Baseline model
m2MLBase <- lm(TARGET ~ ., data = m3train)
knitr::kable(summary(m2MLBase)$coefficients, digits = 3L,
             caption = 'Multiple Regression Model #2 - Base Output')
```

Most of the variables in the dataset are significant, but we can immediately see
a few that are above the 0.05 p-value threshold.

#### Enhanced Model
Backwards stepwise regression is performed and the result is a model with the 
following variables *removed*: `FixedAcidity`, `CitricAcid`, `ResidualSugar`, 
and `Density`.

```{r m2MLTrain}
# ML2: Enhanced model
m2MLFit <- lm(TARGET ~ VolatileAcidity + Chlorides + FreeSulfurDioxide 
              + TotalSulfurDioxide + pH + Sulphates + Alcohol + LabelAppeal 
              + AcidIndex + STARS, data = m3train)
knitr::kable(summary(m2MLFit)$coefficients, digits = 3L,
             caption = 'Multiple Regression Model #2 - Final Output')
```

#### Coefficient Discussion
The intercept of the final model tells us that there are typically ~3 cases of
wine purchased. Some things to note:

  * As expected, the number of stars has the largest, positive impact on the
  number of cases sold. This makes sense -- the higher the rating of a wine, the
  more of it we would expect to be purchased.
  * Label appeal has the second largest impact on the number of cases sold. They
  say one shouldn't judge a book by its cover (or in this case, a wine by its
  label), but it does appear to have quite a significant impact on sales. 
  * Alcohol also has a small positive impact on the number of cases purchased.
  This means that the more alcohol a wine has, the more cases of it will be 
  purchased. 
  * Acid Index, Sulphates, pH, Chlorides, and VolatileAcidity all have a
  negative impact on the number of cases sold.

# SELECT MODELS
## Model Selection Criteria
We will look at RMSE and MAE as our metrics and select the model which performs
best on both. If no model performs best on both, the AIC of the model results on
the training set will be used to break the tie, as the AIC is an estimate of the
average out-of-sample performance of the model.

```{r compTable}
compTable <- data.table(Models = c("Poisson 1", "Poisson 2",
                                   "Neg. Binom. 1", "Neg. Binom. 2",
                                   "MultiLinear 1", "MultiLinear 2"),
                        RMSE = double(6), MAE = double(6),
                        TrainAIC = double(6))
```

## Model Test Results
The test set data needs to be processed identically to the training set. Since
the number of cases needs to be integral the predicted values will be rounded to
the nearest integer.

### Poisson Model 1

```{r m1TestProcess}
# P1 & NB1: process test data
m1tst <- copy(tstSet)
m1tst[, Rated := factor(ifelse(is.na(STARS), 'Unrated', 'Rated'),
                        levels = c('Unrated', 'Rated'),
                        labels = c('Unrated', 'Rated'))]
set.seed(89)
m1tstI <- preProcess(m1tst, method =c('nzv', 'bagImpute'))
m1tst[is.na(STARS), STARS := 0L]
m1tstImp <- predict(m1tstI, newdata = m1tst)
```
```{r m1PTest}
# P1: Test Model
m1PtstPred <- round(predict(m1PFit, newdata = m1tstImp))

compTable[1, 2] <- m1PRMSE <- RMSE(m1PtstPred, m1tstImp$TARGET)
compTable[1, 3] <- m1PMAE <- MAE(m1PtstPred, m1tstImp$TARGET)
compTable[1, 4] <- m1PFitAIC
```

**Poisson Model 1** has an RMSE of 
`r prettyNum(m1PRMSE, digits = 2L)`, an MAE of 
`r prettyNum(m1PMAE, digits = 2L)`, and a training AIC of
`r prettyNum(m1PFitAIC, digits = 2L)`.

### Poisson Model 2

```{r m2TestProcess}
set.seed(525)
m2test <- copy(tstSet[,-1])
m2test[,c(2:8,11:12)] <- abs(m2test[,c(2:8,11:12)])
m2test$STARS[is.na(m2test$STARS)] <- 0
m2test.impute <- mice(m2test, method = 'pmm', print = FALSE)
m2test <- complete(m2test.impute)
m2test.r <- m2test$TARGET
m2test.p <- m2test[,-1]
pre.process <- preProcess(m2test.p, method = c("YeoJohnson"))
m2test.p <- predict(pre.process, m2test.p)
m2test[,c(2:13,15)] <- m2test.p[,c(1:12,14)]
```
```{r m2PTest}
# ZIP2: Test Model
m2PtstPred = round(predict(m2.fpmodel, newdata = m2test))
compTable[2, 2] <- m2PRMSE <- RMSE(m2PtstPred, m2test.r)
compTable[2, 3] <- m2PMAE <- MAE(m2PtstPred, m2test.r)
compTable[2, 4] <- m2.poisson.aic
```

**Poisson Model 2** has an RMSE of 
`r prettyNum(m2PRMSE, digits = 2L)`, an MAE of 
`r prettyNum(m2PMAE, digits = 2L)`, and a training AIC of 
`r prettyNum(m2.poisson.aic, digits = 2L)`.

### Negative Binomial Model 1

```{r m1NBTest}
# NB1: Test Model
m1NBtstPred <- round(predict(m1NBFit, newdata = m1tstImp))
compTable[3, 2] <- m1NBRMSE <- RMSE(m1NBtstPred, m1tstImp$TARGET)
compTable[3, 3] <- m1NBMAE <- MAE(m1NBtstPred, m1tstImp$TARGET)
compTable[3, 4] <- m1NBFitAIC 
```

**Negative Binomial Model 1** has an RMSE of
`r prettyNum(m1NBRMSE, digits = 2L)`, an MAE of
`r prettyNum(m1NBMAE, digits = 2L)`, and a training AIC of
`r prettyNum(m1NBFitAIC, digits = 2L)`.

### Negative Binomial Model 2

```{r m2NBTest}
# NB2: Test Model
m2NBtstPred <- round(predict(m2NBFit, newdata = m3test))
compTable[4, 2] <- m2NBRMSE <- RMSE(m2NBtstPred, m3test$TARGET)
compTable[4, 3] <- m2NBMAE <- MAE(m2NBtstPred, m3test$TARGET)
compTable[4, 4] <- m2NBFitAIC <- AIC(m2NBFit)
```

**Negative Binomial Model 2** has an RMSE of
`r prettyNum(m2NBRMSE, digits = 2L)`, an MAE of
`r prettyNum(m2NBMAE, digits = 2L)`, and a training AIC of
`r prettyNum(m2NBFitAIC, digits = 2L)`.

### Multiple Linear Regression Model 1

```{r m1MLRTest}
# MLR1: Test Model
m1MLRtstPred <- round(predict(m1MLRFit, newdata = m2test))
compTable[5, 2] <- m1MLRRMSE <- RMSE(m1MLRtstPred, m2test.r)
compTable[5, 3] <- m1MLRMAE <- MAE(m1MLRtstPred, m2test.r)
compTable[5, 4] <- m1MLRFitAIC <- AIC(m1MLRFit)
```

**Multiple Linear Regression Model 2** has an RMSE of
`r prettyNum(m1MLRRMSE, digits = 2L)`, an MAE of
`r prettyNum(m1MLRMAE, digits = 2L)`, and a training AIC of
`r prettyNum(m1MLRFitAIC, digits = 2L)`.

### Multiple Linear Regression Model 2

```{r m2MLTest}
# MLR2: Test Model
m2MLtstPred <- round(predict(m2MLFit, newdata = m3test))
compTable[6, 2] <- m2MLMSE <- RMSE(m2MLtstPred, m3test$TARGET)
compTable[6, 3] <- m2MLMAE <- MAE(m2MLtstPred, m3test$TARGET)
compTable[6, 4] <- m2MLFitAIC <- AIC(m2MLFit)
```

**Multiple Linear Regression Model 1** has an RMSE of
`r prettyNum(m2MLMSE, digits = 2L)`, an MAE of
`r prettyNum(m2MLMAE, digits = 2L)`, and a training AIC of
`r prettyNum(m2MLFitAIC, digits = 2L)`.

## Comparsion

```{r modelTestResults}
# Compare results on test set
kable(compTable, digits = 3L,
      caption = "Model Results on Test Set")
```

## Selection

# PREDICTION

# REFERENCES

  * Kuhn, M. (2019, March 27). *The `caret` Package*.
  https://topepo.github.io/caret/pre-processing.html#imputation
  * Yeo, I., & Johnson, R. (2000). A New Family of Power Transformations to
  Improve Normality or Symmetry. **Biometrika**, 87(4), 954--959.
  Retrieved November 26, 2020, from http://www.jstor.org/stable/2673623

# CODE APPENDIX
```{r setup2, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, eval=FALSE)
```
<!-- Assignment -->
```{r loadData}
```
<!-- Data Exploration -->
```{r missingVal}
```
```{r statsN}
```
```{r graphsD}
```
```{r graphsH}
```
```{r corrgram}
```
```{r targetH}
```
<!-- Data Preparation -->
```{r trainTestSplit}
```
```{r model1addVars}
```
```{r model1impute}
```
```{r m1postImputeD}
```
```{r m1postImputeH}
```
```{r m2clean}
```
```{r model3impute}
```
<!-- Build Models -->
```{r m1PTrain}
```
```{r m1PTable}
```
```{r m1PFitCheck}
```
```{r m1PvarImp}
```
```{r m1NBTrain}
```
```{r m1NBTable}
```
```{r m1NBvarImp}
```
```{r m1PNB}
```
```{r m2poisson}
```
```{r m2PTable}
```
```{r m2PFitCheck}
```
```{r m1MLRFit}
```
```{r m1MLRTable}
```
```{r m2NBTrainBase}
```
```{r m2NBTrain}
```
```{r m2MLTrainBase}
```
```{r m2MLTrain}
```
<!-- Select Models -->
```{r compTable}
```
```{r m1TestProcess}
```
```{r m1PTest}
```
```{r m2TestProcess}
```
```{r m2PTest}
```
```{r m1NBTest}
```
```{r m2NBTest}
```
```{r m1MLRTest}
```
```{r m2MLTest}
```
<!-- Remaining Test sets go here -->
```{r modelTestResults}
```